---
title: "Visualization for SIVEP variables"
author: "Raquel Lana and Narimane"
date: "6 de dezembro de 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Aggregating SIVEP data for year to build API maps

- Packages
```{r}
library(foreign)
library('stringr')
library(tidyverse)
library(ggplot2)
library("scales")#para usar nas escalas gráficas- ver depois.
library("data.table")

```

### Loading SIVEP RData
Environment: NN or RL - RL is **F**.
```{r}
envNN = F

if(envNN){
  load("/SIVEP_clean.RData")
}else{load("/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/SIVEP_clean.RData") #my pc
#}else{load("/home/claudia/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/SIVEP_clean.RData") #desktop Fiocruz  
}

```

- See categories and descriptive statistics.
```{r}
summary(df)

```


- Malaria type and TIPO_LAM per year.
```{r}
LAM_M_type <- df %>%
              group_by(ANO, TIPO_LAM) %>%
              count(RES_EXAM)  
              
options(scipen = 10000)


dev.copy(png, "/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Plots_results/MalariaType_DetectionType", width = 800, height = 900, units = "px", pointsize = 12, res = 100)
    
TL <- ggplot(aes(y = n, x = ANO, color = TIPO_LAM), data = LAM_M_type) +
            geom_line(stat = "identity") +
            facet_grid(RES_EXAM ~ .) +
            scale_x_continuous(labels = seq(2003, 2018, 1), breaks = seq(2003, 2018, 1)) +
            labs(x = "Year", y = "Malaria cases", color = "Detection type") 

dev.off()

```


- ID_LVC_ and TIPO_LAM
From paper **WIEFELS et al. 2016**. Accuracy of the malaria epidemiological surveillance system data in the state of Amazonas. Acta Amazonica:
"The follow-up consultation was related to the active/ passive notification. The patient’s follow-up consultation occurs weeks after the confirmation of the malaria infection (World Health Organization 2012). It is a second consultation to certify the healing and is registered as a new notification, which is why it is essential to distinguish it from the first consultation to avoid duplicates. Two variables - the detection type, called TIPO_LAM (1-Passive, 2-Active and 3- Follow-up consultation) and the follow-up consultation, called ID_LVC (1-Yes, 2-No) certify the follow-up consultation. There were 201 inconsistent notifications checked as a follow- up consultation notification (TIPO_LAM=3) but also as its opposite (ID_LVC=2)."
```{r}
LAM_ID_type <- df %>%
              group_by(ANO, TIPO_LAM) %>%
              count(ID_LVC_)  
              

options(scipen = 10000)

dev.copy(png, "/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Plots_results/DetectionType_LVC", width = 800, height = 900, units = "px", pointsize = 12, res = 100)

TLI <- ggplot(aes(y = n, x = ANO, color = TIPO_LAM), data = LAM_ID_type) +
            geom_line(stat = "identity") +
            facet_grid(ID_LVC_ ~ .) +
            scale_x_continuous(labels = seq(2003,2018,1), breaks = seq(2003,2018,1)) +
            labs(x = "Year", y = "Malaria cases stratified by cure verification", color = "Detection type")
        
dev.off()

```

- Malaria type and age
```{r}
M_age_gender <- df %>%
                  group_by(ANO, RES_EXAM) %>%
                  count(AGE_CAT, SEXO)

options(scipen = 10000)

dev.copy(png, "/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Plots_results/AgeCategories_MalariaType", width = 1100, height = 1000, units = "px", pointsize = 12, res = 100)

MA <- ggplot(aes(y = n, x = AGE_CAT, fill = RES_EXAM), data = M_age_gender[which(M_age_gender$SEXO == "F" | M_age_gender$SEXO == "M"),]) +
            geom_bar(stat = "identity", position = "dodge") +
            facet_wrap(~ANO) +
            labs(x = "Age Categories", y = "Malaria cases") +
            theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
            guides(fill = guide_legend(title= "Malaria type")) 

dev.off()

```

- GENDER
```{r}
dev.copy(png, "/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Plots_results/Gender_MalariaType", width = 1100, height = 1000, units = "px", pointsize = 12, res = 100)

MG <- ggplot(aes(y = n, x = SEXO, fill = RES_EXAM), data = M_age_gender[which(M_age_gender$SEXO == "F" | M_age_gender$SEXO == "M"),]) +
            geom_bar(stat = "identity", position = "dodge") +
            facet_wrap(~ANO) +
            labs(x = "Gender", y = "Malaria cases") +
            scale_x_discrete(breaks = c("F", "M"), labels = c("Female", "Male")) +
            guides(fill = guide_legend(title = "Malaria type")) 

dev.off()

```

- To do age and gender stratified by type. Here I summed V+F with F and V. 
```{r}
df1 <- df %>%
        select(RES_EXAM, TIPO_LAM, SEXO, SINTOMAS, ID_LVC_, ANO, AGE_CAT) %>%
        group_by(ANO, TIPO_LAM, SEXO, SINTOMAS, ID_LVC_, AGE_CAT) %>%
        count(RES_EXAM) %>%
        spread(RES_EXAM, n, fill = 0) %>%
        rename(FALCI = "Falciparum") %>%
        rename(VF = "V+F") %>%
        rename(VIVAX = "Vivax") %>%
        mutate(Falciparum = FALCI + VF) %>%
        mutate(Vivax = VIVAX + VF) %>%
        select(ANO, TIPO_LAM, SEXO, SINTOMAS, ID_LVC_, AGE_CAT, Falciparum, Vivax) %>%
        gather(key = 'MALARIA_TYPE', value = 'CASES', -c(ANO, TIPO_LAM, SEXO, SINTOMAS, ID_LVC_, AGE_CAT))

```


- AGE and GENDER stratified by malaria type
```{r}
#Change the title when change the type
dev.copy(png, "/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Plots_results/AgeCategory_Gender_Falciparum", width = 1100, height = 1000, units = "px", pointsize = 12, res = 100)

#Change the type: Falciparum/Vivax
df2 <- df1 %>% filter(MALARIA_TYPE == "Falciparum")

AG <- ggplot(aes(y = CASES, x = AGE_CAT, fill = SEXO), data = df2[which(df2$SEXO == "F" | df2$SEXO == "M"),]) +
            geom_bar(stat = "identity", position = "dodge") +
            facet_wrap(~ANO) +
            labs(x = "Age categories", y = "Malaria falciparum cases") + #Change to vivax/falciparum
            theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
            guides(fill = guide_legend(title = "Gender")) 

dev.off()

```


- SINTOMAS- ver com LVC e com TIPO_LAM
```{r}
#Change the title when change the type falciparum/vivax
dev.copy(png, "/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Plots_results/Symptoms_DetectionType_Vivax", width = 1100, height = 1000, units = "px", pointsize = 12, res = 100)

# SCD <- ggplot(aes(y = CASES, x = SINTOMAS, fill = ID_LVC_), data = df1[df1$MALARIA_TYPE == "Vivax",]) +
#             geom_bar(stat = "identity", position = "dodge") +
#             facet_wrap(~ANO) +
#             labs(x = "Symptoms", y = "Malaria vivax cases") +
#             theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
#             guides(fill = guide_legend(title ="Cure Verification"))
# dev.off()

SCD2 <- ggplot(aes(y = CASES, x = SINTOMAS, fill = TIPO_LAM), data = df1[df1$MALARIA_TYPE == "Vivax",]) +
            geom_bar(stat = "identity", position = "dodge") +
            facet_wrap(~ANO) +
            labs(x = "Symptoms", y = "Malaria vivax cases") +
            theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
            guides(fill = guide_legend(title ="Detection type"))
dev.off()

# SCD3 <- ggplot(aes(y = CASES, x = TIPO_LAM, fill = SINTOMAS), data = df1[df1$MALARIA_TYPE == "Falciparum",]) +
#             geom_bar(stat = "identity", position = "dodge") +
#             facet_wrap(~ID_LVC_) +
#             labs(x = "Detection type stratified by Cure verification", y = "Malaria falciparum cases") +
#             guides(fill = guide_legend(title ="Symptoms"))
# 
# dev.off()
            
```

Pregnant - check percentage of pregnant
```{r}
unique(df$GESTANTE_)
preg <- df %>%
          group_by(ANO, RES_EXAM, SEXO) %>%
          count(GESTANTE_)

pms <- ggplot(aes(y = n, x = GESTANTE_), data = preg) +
            geom_bar(stat = "identity", position = "dodge") +
            facet_wrap(~ANO) 
         
  
```


### API by gender and age categories
- I have to reorder the categories for AGE_CAT again.
```{r}
API_strat <- read.csv("/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Malaria_Mapping_TimeSeries_Data/API_admlevels_year_STRAT.csv") #pc path

API_strat <- read.csv("/home/claudia/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Malaria_Mapping_TimeSeries_Data/API_admlevels_year_STRAT.csv") # Fiocruz desktop


head(API_strat)

cat_order <- c("0-4", "5-9", "10-14", "15-19", "20-24", "25-29", "30-34", "35-39", "40-44", "45-49", "50-54", "55-59", "60-64", "65-69", "70-74", "75-79", "80+")

API_strat <- API_strat %>% 
                        mutate(AGE_CAT1 = factor(AGE_CAT, levels = cat_order)) 

```

### API stratified by age and gender per malaria type and adm levels. 
- Brazil
```{r}
APIfb <- API_strat %>% filter(TYPE == "Falciparum" & CODE == 1)

dev.copy(png, "/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Plots_results/API_Brazil_falciparum.png", width = 900, height = 800, units = "px", pointsize = 12, res = 100)

dev.copy(png, "/home/claudia/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Plots_results/API_Brazil_falciparum.png", width = 900, height = 800, units = "px", pointsize = 12, res = 100)

bf <- ggplot(aes(y = API, x = AGE_CAT1, fill = GENDER), data = APIfb) +
            geom_bar(stat = "identity", position = "dodge") +
            facet_wrap(~YEAR) +
            labs(x = "Age categories", y = "Malaria falciparum API in Brazil") + 
            theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
            guides(fill = guide_legend(title = "Gender")) 
dev.off()

APIvb <- API_strat %>% filter(TYPE == "Vivax" & CODE == 1)

dev.copy(png, "/home/claudia/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Plots_results/API_Brazil_vivax.png", width = 900, height = 800, units = "px", pointsize = 12, res = 100)

bv <- ggplot(aes(y = API, x = AGE_CAT1, fill = GENDER), data = APIvb) +
            geom_bar(stat = "identity", position = "dodge") +
            facet_wrap(~YEAR) +
            labs(x = "Age categories", y = "Malaria vivax API in Brazil") + 
            theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
            guides(fill = guide_legend(title = "Gender")) 

dev.off()

```

- State  
Codes: -feito: 11     12     13-     14     15     16     17     21     51
```{r}
state <- 13
APIfs <- API_strat %>% filter(TYPE == "Falciparum" & CODE == state)

#dev.copy(png, "/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Plots_results/API/API_AC_vivax", width = 900, height = 800, units = "px", pointsize = 12, res = 100)

dev.copy(png, "/home/claudia/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Plots_results/API/API_AM_falciparum.png", width = 900, height = 800, units = "px", pointsize = 12, res = 100)

sf <- ggplot(aes(y = API, x = AGE_CAT1, fill = GENDER), data = APIfs) +
            geom_bar(stat = "identity", position = "dodge") +
            facet_wrap(~YEAR) +
            labs(x = "Age categories", y = "Malaria falciparum API", title = "Amazonas") + 
            theme(axis.text.x = element_text(angle = 90, hjust = 1), plot.title = element_text(hjust = 0.5)) +
            guides(fill = guide_legend(title = "Gender")) 

dev.off()

APIvs <- API_strat %>% filter(TYPE == "Vivax" & CODE == state)

dev.copy(png, "/home/claudia/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Plots_results/API/API_AM_vivax.png", width = 900, height = 800, units = "px", pointsize = 12, res = 100)

sv <- ggplot(aes(y = API, x = AGE_CAT1, fill = GENDER), data = APIvs) +
            geom_bar(stat = "identity", position = "dodge") +
            facet_wrap(~YEAR) +
            labs(x = "Age categories", y = "Malaria vivax API", title = "Amazonas") + 
            theme(axis.text.x = element_text(angle = 90, hjust = 1), plot.title = element_text(hjust = 0.5)) +
            guides(fill = guide_legend(title = "Gender")) 

dev.off()

```

### Occupation and Gender
- 21.8% missings
- **COD_OCUP**: code of the main activity performed by the patient in the last 15 days.   
1- Agriculture  
2- Livestock  
3- Domestic  
4- Turism  
5 Artisanal and small mining  
6 Exploration plant  
7 Hunting and fishing  
8 Construction of Roads and Dams  
9 Mining  
10 Traveler  
11 Others  
99 Ignored: **Men: 57.69%, AGE_CAT 0-4 and 5-9: 57.3%, AGE_CAT 0-4, 5-9 and 10-14: 71.8%**  
NA's:   

```{r}
summary(df$COD_OCUP)

missings <- (672912 + 581937)/nrow(df) * 100 

df %>% filter(is.na(COD_OCUP) & AGE_CAT == "0-4")  %>% head(5)

df$COD_OCUP1 <- factor(df$COD_OCUP, c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11"), c("Agriculture", "Livestock", "Domestic", "Turism", "Artisanal and small mining", "Exploration plant",  "Hunting and fishing", "Construction of Roads and Dams", "Mining", "Traveler", "Others"))

summary(df$COD_OCUP1)


```

Plot
```{r}
Occup_gender <- df %>%
                  group_by(ANO, SEXO, RES_EXAM) %>%
                  count(COD_OCUP1) %>%
                  spread(RES_EXAM, n, fill = 0) %>%
                  rename(FALCI = "Falciparum") %>%
                  rename(VF = "V+F") %>%
                  rename(VIVAX = "Vivax") %>%
                  mutate(Falciparum = FALCI + VF) %>%
                  mutate(Vivax = VIVAX + VF) %>%
                  select(ANO, SEXO, COD_OCUP1, Falciparum, Vivax) %>%
                  gather(key = 'MALARIA_TYPE', value = 'CASES', -c(ANO, SEXO, COD_OCUP1)) 
            

Occup_gender1 <- Occup_gender %>% filter(MALARIA_TYPE == "Falciparum") %>% filter(SEXO == "F" | SEXO == "M")


#dev.copy(png, "/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Plots_results/Occup_Gender_vivax_Brazil.png", width = 1000, height = 900, units = "px", pointsize = 12, res = 110)

dev.copy(png, "/home/claudia/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Plots_results/Occup_Gender_falciparum_Brazil", width = 900, height = 800, units = "px", pointsize = 12, res = 100)

OG <- ggplot(aes(y = CASES, x = SEXO, fill = COD_OCUP1), data =  Occup_gender1) +
            geom_bar(stat = "identity", position = "dodge") +
            facet_wrap(~ANO) +
            labs(x = "Gender", y = "Falciparum malaria cases") +
            scale_x_discrete(breaks = c("F", "M"), labels = c("Female", "Male")) +
            guides(fill = guide_legend(title = "Occupation")) 

dev.off()


```


### Differecen between date of 1st symptoms, date of exam and date of start of treatment
- DT_SINTO: 337504 NA's: 5.86%
- DT_EXAME: 225 NA's: 0.0039%
- DT_TRATA: 274199 NA's: 4.76%  

Obs.: Ask if treatment can begin before the examination has been done. Ask if there are any criteria to do it if it is positive.
The examination was done before the first symptoms is not a problem, it may be related to active detection. Check this out. See also symptomatic when this happens. What would be the reasonable amount of difference days to exclude from the analysis? Something that could refer to error of completion.
```{r}
summary(df$DT_SINTO)
summary(df$DT_EXAME)
summary(df$DT_TRATA)

Date_diff <- df %>%
              mutate(DIF.SYM.EXAM = as.numeric(DT_EXAME - DT_SINTO),
                     DIF.EXAM.TREAT = as.numeric(DT_TRATA - DT_EXAME),
                     DIF.SYM.TREAT = as.numeric(DT_TRATA - DT_SINTO))
            
summary(Date_diff$DIF.SYM.EXAM)
summary(Date_diff$DIF.EXAM.TREAT)
summary(Date_diff$DIF.SYM.TREAT)

Date_diff %>% filter(DIF.SYM.EXAM < 0) %>% count() # 14515
Date_diff %>% filter(DIF.SYM.EXAM > 3) %>% count() # > 3rdQ: 1329176
Date_diff %>% filter(DIF.SYM.EXAM > 100) %>% count() # 17773
Date_diff %>% filter(DIF.SYM.EXAM > 500) %>% count() # 5249

Date_diff %>% filter(DIF.EXAM.TREAT < 0) %>% count() # 35056
Date_diff %>% filter(DIF.EXAM.TREAT > 0) %>% count() # 121571
Date_diff %>% filter(DIF.EXAM.TREAT > 100) %>% count() # 6185
Date_diff %>% filter(DIF.EXAM.TREAT > 500) %>% count() # 3846

Date_diff %>% filter(DIF.SYM.TREAT < 0) %>% count() # 348
Date_diff %>% filter(DIF.SYM.TREAT > 0) %>% count() # 4149692
Date_diff %>% filter(DIF.SYM.TREAT > 100) %>% count() # 18132
Date_diff %>% filter(DIF.SYM.TREAT > 500) %>% count() # 7542

```

### Symptomatic: if have symptoms before the exam  
- If the difference between date of exam and 1st symptoms is < 0 and have or not symptoms.
- Check symptoms by missings in date of symptoms, exame and treatment.
```{r}
Date_diff %>%
        filter(is.na(DT_SINTO)) %>%
        count(SINTOMAS)

Date_diff %>%
        filter(is.na(DT_EXAME)) %>%
        count(SINTOMAS)

Date_diff %>%
        filter(is.na(DT_TRATA)) %>%
        count(SINTOMAS)


sym_exam <- Date_diff %>%
                      mutate(DIF.SYM.EXAM.CAT = cut(DIF.SYM.EXAM, 
                                                    breaks = c(-Inf, 0, Inf), 
                                                    labels = c("<=0", ">0")),
                             DIF.EXAM.TREAT.CAT = cut(DIF.EXAM.TREAT, 
                                                    breaks = c(-Inf, 0, Inf), 
                                                    labels = c("<=0", ">0")),
                             DIF.SYM.TREAT.CAT = cut(DIF.SYM.TREAT, 
                                                    breaks = c(-Inf, 0, Inf), 
                                                    labels = c("<=0", ">0")),
                           ) %>%
                        group_by(SINTOMAS) %>%
                        count(DIF.SYM.EXAM.CAT)
#                        count(DIF.EXAM.TREAT.CAT)
#                        count(DIF.SYM.TREAT.CAT)


```



