---
title: "Visualization for SIVEP variables"
author: "Raquel Lana and Narimane"
date: "6 de dezembro de 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Aggregating SIVEP data for year to build API maps

- Packages
```{r}
library(foreign)
library('stringr')
library(tidyverse)
library(ggplot2)
library("scales")#para usar nas escalas gráficas- ver depois.
library("data.table")

```

### Loading SIVEP RData
Environment: NN or RL - RL is **F**.
```{r}
envNN = F

if(envNN){
  load("/SIVEP_clean.RData")
}else{load("/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/SIVEP_clean.RData")
}

```

- See categories and descriptive statistics.
```{r}
summary(df)

```


- Malaria type and TIPO_LAM per year.
```{r}
LAM_M_type <- df %>%
              group_by(ANO, TIPO_LAM) %>%
              count(RES_EXAM)  
              
options(scipen = 10000)


dev.copy(png, "/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Plots_results/MalariaType_DetectionType", width = 800, height = 900, units = "px", pointsize = 12, res = 100)
    
TL <- ggplot(aes(y = n, x = ANO, color = TIPO_LAM), data = LAM_M_type) +
            geom_line(stat = "identity") +
            facet_grid(RES_EXAM ~ .) +
            scale_x_continuous(labels = seq(2003, 2018, 1), breaks = seq(2003, 2018, 1)) +
            labs(x = "Year", y = "Malaria cases", color = "Detection type") 

dev.off()

```


- ID_LVC_ and TIPO_LAM
From paper **WIEFELS et al. 2016**. Accuracy of the malaria epidemiological surveillance system data in the state of Amazonas. Acta Amazonica:
"The follow-up consultation was related to the active/ passive notification. The patient’s follow-up consultation occurs weeks after the confirmation of the malaria infection (World Health Organization 2012). It is a second consultation to certify the healing and is registered as a new notification, which is why it is essential to distinguish it from the first consultation to avoid duplicates. Two variables - the detection type, called TIPO_LAM (1-Passive, 2-Active and 3- Follow-up consultation) and the follow-up consultation, called ID_LVC (1-Yes, 2-No) certify the follow-up consultation. There were 201 inconsistent notifications checked as a follow- up consultation notification (TIPO_LAM=3) but also as its opposite (ID_LVC=2)."
```{r}
LAM_ID_type <- df %>%
              group_by(ANO, TIPO_LAM) %>%
              count(ID_LVC_)  
              

options(scipen = 10000)

dev.copy(png, "/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Plots_results/DetectionType_LVC", width = 800, height = 900, units = "px", pointsize = 12, res = 100)

TLI <- ggplot(aes(y = n, x = ANO, color = TIPO_LAM), data = LAM_ID_type) +
            geom_line(stat = "identity") +
            facet_grid(ID_LVC_ ~ .) +
            scale_x_continuous(labels = seq(2003,2018,1), breaks = seq(2003,2018,1)) +
            labs(x = "Year", y = "Malaria cases stratified by cure verification", color = "Detection type")
        
dev.off()

```

- Malaria type and age
```{r}
M_age_gender <- df %>%
                  group_by(ANO, RES_EXAM) %>%
                  count(AGE_CAT, SEXO)

options(scipen = 10000)

dev.copy(png, "/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Plots_results/AgeCategories_MalariaType", width = 1100, height = 1000, units = "px", pointsize = 12, res = 100)

MA <- ggplot(aes(y = n, x = AGE_CAT, fill = RES_EXAM), data = M_age_gender[which(M_age_gender$SEXO == "F" | M_age_gender$SEXO == "M"),]) +
            geom_bar(stat = "identity", position = "dodge") +
            facet_wrap(~ANO) +
            labs(x = "Age Categories", y = "Malaria cases") +
            theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
            guides(fill = guide_legend(title= "Malaria type")) 

dev.off()

```

- GENDER
```{r}
dev.copy(png, "/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Plots_results/Gender_MalariaType", width = 1100, height = 1000, units = "px", pointsize = 12, res = 100)

MG <- ggplot(aes(y = n, x = SEXO, fill = RES_EXAM), data = M_age_gender[which(M_age_gender$SEXO == "F" | M_age_gender$SEXO == "M"),]) +
            geom_bar(stat = "identity", position = "dodge") +
            facet_wrap(~ANO) +
            labs(x = "Gender", y = "Malaria cases") +
            scale_x_discrete(breaks = c("F", "M"), labels = c("Female", "Male")) +
            guides(fill = guide_legend(title = "Malaria type")) 

dev.off()

```

- To do age and gender stratified by type. Here I summed V+F with F and V. 
```{r}
df1 <- df %>%
        select(RES_EXAM, TIPO_LAM, SEXO, SINTOMAS, ID_LVC_, ANO, AGE_CAT) %>%
        group_by(ANO, TIPO_LAM, SEXO, SINTOMAS, ID_LVC_, AGE_CAT) %>%
        count(RES_EXAM) %>%
        spread(RES_EXAM, n, fill = 0) %>%
        rename(FALCI = "Falciparum") %>%
        rename(VF = "V+F") %>%
        rename(VIVAX = "Vivax") %>%
        mutate(Falciparum = FALCI + VF) %>%
        mutate(Vivax = VIVAX + VF) %>%
        select(ANO, TIPO_LAM, SEXO, SINTOMAS, ID_LVC_, AGE_CAT, Falciparum, Vivax) %>%
        gather(key = 'MALARIA_TYPE', value = 'CASES', -c(ANO, TIPO_LAM, SEXO, SINTOMAS, ID_LVC_, AGE_CAT))

```


- AGE and GENDER stratified by malaria type
```{r}
#Change the title when change the type
dev.copy(png, "/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Plots_results/AgeCategory_Gender_Falciparum", width = 1100, height = 1000, units = "px", pointsize = 12, res = 100)

#Change the type: Falciparum/Vivax
df2 <- df1 %>% filter(MALARIA_TYPE == "Falciparum")

AG <- ggplot(aes(y = CASES, x = AGE_CAT, fill = SEXO), data = df2[which(df2$SEXO == "F" | df2$SEXO == "M"),]) +
            geom_bar(stat = "identity", position = "dodge") +
            facet_wrap(~ANO) +
            labs(x = "Age categories", y = "Malaria falciparum cases") + #Change to vivax/falciparum
            theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
            guides(fill = guide_legend(title = "Gender")) 

dev.off()

```


- SINTOMAS- ver com LVC e com TIPO_LAM
```{r}
#Change the title when change the type falciparum/vivax
dev.copy(png, "/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Plots_results/Symptoms_DetectionType_Vivax", width = 1100, height = 1000, units = "px", pointsize = 12, res = 100)

# SCD <- ggplot(aes(y = CASES, x = SINTOMAS, fill = ID_LVC_), data = df1[df1$MALARIA_TYPE == "Vivax",]) +
#             geom_bar(stat = "identity", position = "dodge") +
#             facet_wrap(~ANO) +
#             labs(x = "Symptoms", y = "Malaria vivax cases") +
#             theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
#             guides(fill = guide_legend(title ="Cure Verification"))
# dev.off()

SCD2 <- ggplot(aes(y = CASES, x = SINTOMAS, fill = TIPO_LAM), data = df1[df1$MALARIA_TYPE == "Vivax",]) +
            geom_bar(stat = "identity", position = "dodge") +
            facet_wrap(~ANO) +
            labs(x = "Symptoms", y = "Malaria vivax cases") +
            theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
            guides(fill = guide_legend(title ="Detection type"))
dev.off()

# SCD3 <- ggplot(aes(y = CASES, x = TIPO_LAM, fill = SINTOMAS), data = df1[df1$MALARIA_TYPE == "Falciparum",]) +
#             geom_bar(stat = "identity", position = "dodge") +
#             facet_wrap(~ID_LVC_) +
#             labs(x = "Detection type stratified by Cure verification", y = "Malaria falciparum cases") +
#             guides(fill = guide_legend(title ="Symptoms"))
# 
# dev.off()
            
```

Pregnant - check percentage of pregnant
```{r}
unique(df$GESTANTE_)
preg <- df %>%
          group_by(ANO, RES_EXAM, SEXO) %>%
          count(GESTANTE_)

pms <- ggplot(aes(y = n, x = GESTANTE_), data = preg) +
            geom_bar(stat = "identity", position = "dodge") +
            facet_wrap(~ANO) 
         
  
```








