---
title: "Malaria_Cases_API_final_functions"
author: "Raquel Lana and Narimane Nekkab"
date: "4 de fevereiro de 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Daniel, a ideia desse código é que ele tenha apenas as funções, mas como ainda estou fazendo, vc verá ao longo do scrpti as partes que devem ser emovidas para um "header file". Se vc ver trechos de códigos que podem ser melhorados, por favor, fique a vontade. Só peço rpa comentar para eu veja o que mudou para aprender a fazer do melhor jeito. Eu comentei tb nas figuras, onde não consegui salvar ainda. Se puder ver isso! Até consigo gerar as figuras, mas sem salvar. 


# Final functions for the analyzes and figures of the first paper

- Packages - move to header file together the parameters.
```{r}
library("ggplot2")
library("forecast")
library("tseries")
library("stringr")
library("gridExtra")
library("data.table")
library("fUnitRoots")
library("FitAR")
library("lubridate")
#library("parallel")
library("TSA")
library("dplyr")
library("foreach")
library("tstools")
library("ggfortify")
library("RColorBrewer")
library("colorspace")
library("foreign")
library("tidyverse")
library("maptools")
library("rgdal")

```

- mover para o header file para rodar a função abaixo
```{r}
# Get SIVEP data cleaned up and by species
#SIVEPFilePath = "~/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/SIVEP_clean.RData"
SIVEPFilePath = "/home/rlana/Dados/SIVEP_clean_2003_2018.RData"
#SIVEPFilePath = "/home/daniel/temp/Malaria_mapping"
StartYear = 2004 # (inferior limit = 2003)
EndYear = 2018 # (superior limit = 2018)
AgTimeUnit = "week" # ("week", "month")
byNotification = T 
byResidence = F
byInfection = F
byPCD = F
byPCDACD = T

```

### Function to organize the SIVEP database per:
- DATE: from the date of notification being that it can provide the first day of the epidemiological week in which the notification date was made or the month of this date.    
- LEVEL: Country (BR), State (UF), Municipality (MU).   
- CODE: Country (1), State (Amazon Basin states IBGE codes) and Municipality (Amazon Basin municipalities IBGE codes).  
- TYPE: Falciparum ou Vivax malaria
- Cases: number of cases counted by date, level, code and type of malaria.  
```{r}
getWEEKLY_SIVEP_MALARIA_TYPE <- function(SIVEPFilePath, StartYear, EndYear, AgTimeUnit){
  
  # Get SIVEP raw notification data
  load(SIVEPFilePath)

  # Choose time period
  df <- df[which(df[,"DT_NOTIF"] >= paste0(StartYear,"-01-01") & df[,"DT_NOTIF"] <= paste0(EndYear, "-12-31")),] # parece não funcionar para residencia e prob infe, mas na sequencia corta...quando vai para merge com população
  
  
  # Replace missing state code with municipality state code
  df <- df %>% 
    mutate(UF_NOTIF = as.integer(substr(as.character(MUN_NOTI), 1, 2)),
           UF_RESID = as.integer(substr(as.character(MUN_RESI), 1, 2)),
           UF_INFEC = as.integer(substr(as.character(MUN_INFE), 1, 2)))
  
  # Select administrative level via code
  if(byNotification){
    df$PAIS_CODE <- 1
    names(df)[names(df) == 'UF_NOTIF'] <- 'UF_CODE'
    names(df)[names(df) == 'MUN_NOTI'] <- 'MUN_CODE'
  }
  if(byResidence){
    names(df)[names(df) == 'PAIS_RES'] <- 'PAIS_CODE'
    names(df)[names(df) == 'UF_RESID'] <- 'UF_CODE'
    names(df)[names(df) == 'MUN_RESI'] <- 'MUN_CODE'
  }
  if(byInfection){
    names(df)[names(df) == 'PAIS_INF'] <- 'PAIS_CODE'
    names(df)[names(df) == 'UF_INFEC'] <- 'UF_CODE'
    names(df)[names(df) == 'MUN_INFE'] <- 'MUN_CODE'
  }
  
  #Select the cases by Type of Detection
  if(byPCD){
    df <- df %>%
      filter(TIPO_LAM == "PCD")
  }

  if(byPCDACD){
    df <- df %>%
      filter(TIPO_LAM == "PCD" | TIPO_LAM == "ACD")
  }
  
  # Split data by malaria type and aggregate to MU - level
  week_malaria_type_muni <- df %>%
    group_by(DT_NOTIF, MUN_CODE) %>%
    count(RES_EXAM) %>%
    mutate(LEVEL = "MU") %>%
    select(DT_NOTIF, LEVEL, MUN_CODE, RES_EXAM, n) %>%
    spread(RES_EXAM, n, fill = 0) %>%
    rename(CODE = MUN_CODE,
           FALCI = "F",
           VF = "V+F",
           VIVAX = "V") %>%
    mutate(Falciparum = FALCI + VF,
           Vivax = VIVAX + VF,
           DATE = floor_date(as.Date(DT_NOTIF, format = "%Y-%m-%d"), unit = AgTimeUnit)
           ) %>%
    ungroup %>%
    select(-DT_NOTIF) %>%
    select(DATE, LEVEL, CODE, Falciparum, Vivax) %>%
    gather(key = TYPE, value = CASES, -c(DATE, LEVEL, CODE))

  
  # UF - level
  week_malaria_type_state <- df %>%
    group_by(DT_NOTIF, UF_CODE) %>%
    count(RES_EXAM) %>%
    mutate(LEVEL = "UF") %>%
    select(DT_NOTIF, LEVEL, UF_CODE, RES_EXAM, n) %>%
    spread(RES_EXAM, n, fill = 0) %>%
    rename(CODE = UF_CODE,
           FALCI = "F",
           VF = "V+F",
           VIVAX = "V") %>%
    mutate(Falciparum = FALCI + VF,
           Vivax = VIVAX + VF,
           DATE = floor_date(as.Date(DT_NOTIF, format = "%Y-%m-%d"), unit = AgTimeUnit)
           ) %>%      
    ungroup %>%
    select(-DT_NOTIF) %>%
    select(DATE, LEVEL, CODE, Falciparum, Vivax) %>%
    gather(key = TYPE, value = CASES, -c(DATE, LEVEL, CODE))
  
  # BR - level
  week_malaria_type_brazil <- df %>%
    group_by(DT_NOTIF, PAIS_CODE) %>%
    count(RES_EXAM) %>%
    mutate(LEVEL = "BR")  %>% 
    select(DT_NOTIF, LEVEL, PAIS_CODE, RES_EXAM, n) %>%
    spread(RES_EXAM, n, fill = 0) %>%
    rename(CODE = PAIS_CODE,
           FALCI = "F",
           VF = "V+F",
           VIVAX = "V") %>%
    mutate(Falciparum = FALCI + VF,
           Vivax = VIVAX + VF,
           DATE = floor_date(as.Date(DT_NOTIF, format = "%Y-%m-%d"), unit = AgTimeUnit)
           ) %>%
    ungroup %>%
    select(-DT_NOTIF) %>%
    select(DATE, LEVEL, CODE, Falciparum, Vivax) %>%
    gather(key = TYPE, value = CASES, -c(DATE, LEVEL, CODE))
    
  # Merge together
  SIVEP_MALARIA_TYPE <- rbind(week_malaria_type_brazil, week_malaria_type_state, week_malaria_type_muni)
  rm(df, week_malaria_type_brazil, week_malaria_type_state, week_malaria_type_muni)


  return(SIVEP_MALARIA_TYPE)
  
}


```

Move to header file
```{r}
d <- getWEEKLY_SIVEP_MALARIA_TYPE(SIVEPFilePath, StartYear, EndYear, AgTimeUnit)

```



### Function to agreggate the population size from IBGE and calculate API.
- Move to header file 
```{r}
#IBGEFilePath = "~/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Malaria_Mapping_TimeSeries_Data/BRA_POP_EST.csv"
#IBGEFilePath = "~/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Malaria_Mapping_TimeSeries_Data/BRA_POP_EST.csv"
IBGEFilePath = "/home/rlana/Dados/BRA_POP_EST.csv"
IBGEFilePath = "/home/daniel/temp/Malaria_mapping"
#StartYear = 2004 # (inferior limit = 2003)
#EndYear = 2017 # (superior limit = 2018)
#AgTimeUnit = "week" # ("week", "month")
unitTime <- ifelse(AgTimeUnit == "week", 52, 12) # If AgTimeUnit == "week", unitTime == 52

```

- NAME: name of the administrative levels.  
- YEAR: year of the notification and population estimated.  
- POP_SIZE: population estimated by year and administrative levels
- API: annual parasite index - ver se tem que multiplicar por semanas pois se chama annual p. in.
```{r}
getWEEKLY_SIVEP_API <- function(IBGEFilePath, StartYear, EndYear, AgTimeUnit){
  
  # Get IBGE population size 
  d1 <- read.csv(IBGEFilePath, stringsAsFactors = F)
  # Get names
  ADMIN_NAMES <- read.table("/home/rlana/Dados/BRA_ADMIN_NAMES.csv", sep = " ")
  #ADMIN_NAMES <- read.table("/home/daniel/temp/Malaria_mapping/BRA_ADMIN_NAMES.csv", sep = " ") 
  
  # Reshape the population size table to join with cases table
  d1 <- d1 %>%
    select(-GR) %>%
    filter(!LEVEL %in% c("GR")) %>%
    mutate(CODE = replace(CODE, CODE == 0, 1)) %>%
    gather(key = YEAR, value = POP_SIZE, -c(LEVEL, CODE, NAME)) %>%
    mutate(YEAR = as.integer(str_sub(YEAR, start = 2, end = 5))) %>%
    filter(YEAR >= StartYear & YEAR <= EndYear) 

  # Cut the database only for Amazon Basin States and municipalities
  ADMIN_NAMES <- ADMIN_NAMES %>%
    ungroup %>%
    filter(Level != "GR") %>%
    mutate(CODE = replace(Code, Code == 0, 1))

  d1$STATE = ADMIN_NAMES[match(d1$CODE, ADMIN_NAMES$Code),"UF"] 
  d1$NAME = ADMIN_NAMES[match(d1$CODE, ADMIN_NAMES$Code),"Name"]

  d1 <- d1 %>% 
    filter(STATE == "AC"| STATE == "AM" | STATE == "AP" | STATE == "MA" | STATE == "MT" | STATE == "PA" | STATE == "RO" | STATE == "RR" | STATE == "TO")

  # Add BR level again only for Amazon Basin population
  dBR <- d1 %>% 
    filter(LEVEL == "UF") %>% 
    group_by(YEAR) %>% 
    summarise(POP_SIZE = sum(POP_SIZE)) %>% 
    mutate(LEVEL = "BR",
           CODE = 1,
           NAME = "Amazonia Legal",
           STATE = "NA") %>% 
    select(CODE, LEVEL, NAME, YEAR, POP_SIZE, STATE)
  
  d1 <- bind_rows(dBR, d1)
  
  # Agregate cases by time and create the column year
  d <- d %>% 
    group_by(DATE, LEVEL, CODE, TYPE) %>% 
    summarise(CASES = sum(CASES)) %>% 
    mutate(YEAR = year(DATE))
    
  # Join cases and population tables
  API_MALARIA <- left_join(d1, d)  
  
  # Calculate API and reorder columns and rearrange variables
  API_MALARIA <- API_MALARIA %>% 
    mutate(API = CASES/POP_SIZE * 1000 * unitTime) %>%
    select(LEVEL, CODE, NAME,STATE, DATE, YEAR, TYPE, CASES, POP_SIZE, API) %>%
    arrange(DATE, YEAR, CODE)
  
  return(API_MALARIA)
  
}

```

- Move to header file 
```{r}
d2 <- getWEEKLY_SIVEP_API(IBGEFilePath, StartYear, EndYear, AgTimeUnit)

```

- Save
```{r}
save(d2, file = "/home/rlana/Dados/API_Noti_weeklyAPI_state.Rdata")
#save(d2, file = "/home/daniel/temp/Malaria_mapping")
```


# Function to calculate Pf/Pv and Pv/Total  ratios
- Check if here It's necessary to group by YEAR.
- Obs.: for some strange reason **tidyr** needs another index to do the **spread** sometimes.

Function
```{r}  
getWEEKLY_SIVEP_RATIOS <- function(API_MALARIA, StartYear, EndYear, AgTimeUnit){

  # Spreading to facilitate the ratio's calculus
  RATIOS <- d2 %>%
  select(LEVEL, CODE, NAME, DATE, YEAR, TYPE, CASES) %>%
  group_by_at(vars(-CASES)) %>%  # group by everything other than the value column. 
  mutate(row_id = 1:n()) %>%   #rowid_to_column() - ver essa função pra ver se fica igual
  ungroup() %>%  # build group index
  spread(key = TYPE, value = CASES, fill = 0) %>%    # spread
  select(-row_id) %>%  # drop the index
  mutate(PvPf = Falciparum / Vivax,
         PvTOTAL = Vivax / (Falciparum + Vivax)
         ) %>%
  select(LEVEL, CODE, NAME, DATE, YEAR, PvPf, PvTOTAL)  
  
  return(RATIOS)  
  
}

```

Move to header file
```{r}
d3 <- getWEEKLY_SIVEP_RATIOS(API_MALARIA, StartYear, EndYear, AgTimeUnit) 

```



### Cases and API Brazil time series plot 

Esses plots estão funcionando, só não consegui salvar.  
Não consegui cortar o 2019 do gráfico.  

```{r}
#Move to header
# Date type (by day == "Daily" , week == "Weekly", month == "Monthly", year == "Yearly")
Date_Type = "Weekly"
Plot_Folder = "/home/rlana/Plots/TimeSeries/BR/"
#Plot_Folder = "/home/daniel/temp/Malaria_mapping"

#Cases
Cases_TS_BR <- ggplot(data = subset(d2, LEVEL == "BR"), aes(DATE, CASES, color = TYPE)) +
  stat_summary(fun.y = sum, geom = "line") +
  scale_x_date(breaks = "year",
               date_labels = "%Y") +
  scale_color_manual(values = c("#31a354","#3182bd")) +
  labs(title = paste0("P. vivax and P. falciparum ", "cases in Brazil, ", 
                      StartYear, "-", EndYear), x = "Year", y = paste(Date_Type,"cases")) + 
  guides(color = guide_legend(title = "Plasmodium")) +
  theme(panel.grid.minor.x = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position="bottom",
        legend.text = element_text(size = 5),
        legend.title = element_text(size = 6, face = "italic"),
        legend.margin = margin(0,0,0,0), 
        plot.title = element_text(size = 6),
        axis.title = element_text(size = 5),
        axis.text = element_text(size = 5))

Cases_TS_BR

# Save
ggsave(Cases_TS_BR, filename = paste0("P. vivax and P. falciparum cases in Brazil ", StartYear, "-", EndYear, ".png"), width = 1500/300, height = 1000/300, dpi = 300, path = Plot_Folder, units = "in", device = "png")



#API
API_TS_BR <- ggplot(data = subset(d2, LEVEL == "BR"), aes(DATE, API, color = TYPE)) +
  stat_summary(fun.y = sum, geom = "line") +
  scale_x_date(breaks = "year",
               date_labels = "%Y") +
  scale_color_manual(values = c("#31a354","#3182bd")) +
  labs(title = paste0("P. vivax and P. falciparum ", "API in Brazil, ", 
                      StartYear, "-", EndYear), x = "Year", y = paste(Date_Type,"API")) + 
  guides(color = guide_legend(title = "Plasmodium")) +
  theme(panel.grid.minor.x = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position="bottom",
        legend.text = element_text(size = 5),
        legend.title = element_text(size = 6, face = "italic"),
        legend.margin = margin(0,0,0,0), 
        plot.title = element_text(size = 6),
        axis.title = element_text(size = 5),
        axis.text = element_text(size = 5))

API_TS_BR

# Save
ggsave(API_TS_BR, filename = paste0("P. vivax and P. falciparum API in Brazil ", StartYear, "-", EndYear, ".png"), width = 1500/300, height = 1000/300, dpi = 300, path = Plot_Folder, units = "in", device = "png")

```

- Estou achando o plot de casos e IPA para o Brasil muito igual...é isso mesmo? 



# Time series by STATE- one figure with all states.  
- supp file  
- FALTA salvar! E decidir se faremos assim, ou se faremos uma a um como mais abaixo.

```{r}
Cases_Plot_State <- ggplot(data = subset(d2, LEVEL == "UF"), aes(DATE, CASES, color = TYPE)) + ###testar
  stat_summary(fun.y = sum, geom = "line") +
  scale_x_date(breaks = "year",
               date_labels = "%Y") +
  facet_wrap(.~NAME, ncol = 3) +
  scale_color_manual(values = c("#31a354","#3182bd")) +
  labs(title = paste0("P. vivax and P. falciparum ", "cases in Brazil, ", 
                      StartYear, "-", EndYear), x = "Year", y = paste(Date_Type,"cases")) +
  guides(color = guide_legend(title = "Plasmodium")) +
  theme(panel.grid.minor.x = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position="bottom",
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8, face = "italic"),
        legend.margin = margin(0,0,0,0), 
        plot.title = element_text(size = 8),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7))

Cases_Plot_State

# Save
ggsave(Cases_Plot_State, filename = paste0("P. vivax and P. falciparum cases in Brazil per State", StartYear, "-", EndYear, ".png"), width = 2500/300, height = 2000/300, dpi = 300, path = Plot_Folder, units = "in", device = "png")


API_Plot_State <- ggplot(data = subset(d2, LEVEL == "UF"), aes(DATE, API, color = TYPE)) +
  stat_summary(fun.y = sum, geom = "line") +
  scale_x_date(breaks = "year",
               date_labels = "%Y") +
  facet_wrap(.~NAME, ncol = 3) +
  scale_color_manual(values = c("#31a354","#3182bd")) +
  labs(title = paste0("P. vivax and P. falciparum ", "API in Brazil, ", 
                      StartYear, "-", EndYear), x = "Year", y = paste(Date_Type,"API")) +
 guides(color = guide_legend(title = "Plasmodium")) +
  theme(panel.grid.minor.x = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position="bottom",
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8, face = "italic"),
        legend.margin = margin(0,0,0,0), 
        plot.title = element_text(size = 8),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7))

API_Plot_State

# Save
ggsave(API_Plot_State, filename = paste0("P. vivax and P. falciparum API in Brazil per State", StartYear, "-", EndYear, ".png"), width = 2500/300, height = 2000/300, dpi = 300, path = Plot_Folder, units = "in", device = "png")


```


### State plots - each state separetely
- Decidir!!!!!
```{r}
States_sivep = c("AC","AM","AP","MA","MT","PA","RO","RR","TO")
States_code = c(11, 12, 13, 14,     15     16     17     21     51)   ###terminar aqui

UF_Plots <- foreach(i = 1:length(States_sivep)) %do% {
  Plot_Name = paste0("TS_Plot_", States_sivep[i])
  assign(Plot_Name, ggplot(data = subset(d2, LEVEL == "UF" & CODE == States_sivep[i]), 
                           aes(DATE, CASES, color = TYPE)) +
           stat_summary(fun.y = sum, geom = "line") +
           scale_x_date(breaks = "year", 
                        date_labels = "%Y") +
           scale_color_manual(values = c("#31a354","#3182bd")) +
           facet_wrap(~NAME) +
           labs(title = paste0("P. vivax and P. falciparum cases in ", States_sivep[i], 
                               StartYear, "-", EndYear), x = "Year", y = "Number of Cases") + 
           guides(color = guide_legend(title = "Plasmodium")) +
            theme(panel.grid.minor.x = element_blank(),
                  axis.text.x = element_text(angle = 90, hjust = 1),
                  legend.position="bottom",
                  legend.text = element_text(size = 7),
                  legend.title = element_text(size = 8, face = "italic"),
                  legend.margin = margin(0,0,0,0), 
                  plot.title = element_text(size = 8),
                  axis.title = element_text(size = 7),
                  axis.text = element_text(size = 7))
  )

         }

UF_Plots[2]

#### Ver como salvar automático - talvez isso dentro da função?
# Save
ggsave(UF_Plots[i], filename = paste0("P. vivax and P. falciparum API in Brazil per State", StartYear, "-", EndYear, ".png"), width = 2500/300, height = 2000/300, dpi = 300, path = Plot_Folder, units = "in", device = "png")

```

### Municipalities plots for each state

- Não consegui salvar ainda, mas funciona!!!! 
- Aqui teria que ver se faremos mesmo, se vamos colcoar no suplementar. Para alguns estados, os que tem mais municípios, fica ruim.  
- Se colocarmos no nível de município tb no supp file, talvez fosse melhor usar somente os que tem muitos casos..
```{r}
# States
States_sivep = c("AC","AM","AP","MA","MT","PA","RO","RR","TO")

MUN_Plots <- foreach(i = 1:length(States_sivep)) %do% {
  Plot_Name = paste0("TS_Plot_", States_sivep[i])
  assign(Plot_Name, ggplot(data = subset(dAB, LEVEL == "MU" & 
                                           STATE == States_sivep[i]), 
                           aes(DATE, CASES, color = TYPE)) +
           stat_summary(fun.y = sum, geom = "line") +
           scale_x_date(breaks = "year", 
                        date_labels = "%Y") +
           scale_color_manual(values = c("#31a354","#3182bd")) +
           facet_wrap(~NAME) +
           labs(title = paste0("P. vivax and P. falciparum cases in ", States_sivep[i]," state by municipality, ", 
                               StartYear, "-", EndYear), x = "Year", y = "Number of Cases") + 
           guides(color = guide_legend(title = "Plasmodium species")) +
           theme(panel.grid.minor.x = element_blank(),
                 axis.text.x = element_text(angle = 90, hjust = 1)))
  
  
##Testar
 # ggsave(Plot_Name, )
  
}

MUN_Plots[1]


```




### PArei aqui



## Function to organize the cases per age and gender - doing
- um painel com todos os estados- somatório de todos os anos.
- Para suplementar, seria interessante ter uma figura para cada estado mostrando por ano- facet_wrap por ano.

-REVER SE FAZ POR ANO- VER O PLOT.
```{r}
load("/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/SIVEP_clean.RData") # move to header file

getSTRAT_AGE_GENDER <- function(SIVEPFilePath, StartYear, EndYear, AgTimeUnit){

  
  # Brazil  
  day_malaria_type_brazil_age_gender <- df %>%
    group_by(DT_NOTIF, AGE_CAT, SEXO) %>%
    count(RES_EXAM)%>%
    mutate(LEVEL = "BR",
           CODE = "1") %>%
    select(DT_NOTIF, LEVEL, CODE, RES_EXAM, AGE_CAT, SEXO, n) %>%
    spread(RES_EXAM, n, fill = 0) %>%
    rename(GENDER = SEXO, FALCI = "Falciparum", VF = "V+F", VIVAX = "Vivax") %>%
    mutate(Falciparum = FALCI + VF,
           Vivax = VIVAX + VF,
           DATE = floor_date(as.Date(DT_NOTIF, format = "%Y-%m-%d"), unit = AgTimeUnit)
            ) %>%
    ungroup %>%
    select(-DT_NOTIF) %>%
    mutate(GENDER = ifelse(GENDER == "F", "Female", 
                           ifelse(GENDER == "M", "Male", NA))) %>%
    select(DATE, LEVEL, CODE, AGE_CAT, GENDER, Falciparum, Vivax) %>%
    gather(key = 'TYPE', value = 'CASES', -c(DATE, LEVEL, CODE, AGE_CAT, GENDER))
  
  # State
  day_malaria_type_state_age_gender <- df %>%
    group_by(DT_NOTIF, UF_NOTIF, AGE_CAT, SEXO) %>%
    count(RES_EXAM) %>%
    mutate(LEVEL = "UF") %>%
    select(DT_NOTIF, LEVEL, UF_NOTIF, RES_EXAM, AGE_CAT, SEXO, n) %>%
    spread(RES_EXAM, n, fill = 0) %>%
    rename(CODE = UF_NOTIF, GENDER = SEXO, FALCI = "Falciparum", VF = "V+F", VIVAX = "Vivax") %>%
    mutate(Falciparum = FALCI + VF,
           Vivax = VIVAX + VF,
           DATE = floor_date(as.Date(DT_NOTIF, format = "%Y-%m-%d"), unit = AgTimeUnit)
            ) %>%
    ungroup %>%
    select(-DT_NOTIF) %>%
    mutate(GENDER = ifelse(GENDER == "F", "Female", 
                           ifelse(GENDER == "M", "Male", NA))) %>%
    select(DATE, LEVEL, CODE, AGE_CAT, GENDER, Falciparum, Vivax) %>%
    gather(key = 'TYPE', value = 'CASES', -c(DATE, LEVEL, CODE, AGE_CAT, GENDER))
  
  # Municipality
  day_malaria_type_age_gender <- df %>%
    group_by(DT_NOTIF, MUN_NOTI, AGE_CAT, SEXO) %>%
    count(RES_EXAM) %>%
    mutate(LEVEL = "MU") %>%
    select(DT_NOTIF, LEVEL, MUN_NOTI, RES_EXAM, AGE_CAT, SEXO, n) %>%
    spread(RES_EXAM, n, fill = 0) %>%
    rename(CODE = MUN_NOTI, GENDER = SEXO, FALCI = "Falciparum", VF = "V+F", VIVAX = "Vivax") %>%
    mutate(Falciparum = FALCI + VF,
           Vivax = VIVAX + VF,
           DATE = floor_date(as.Date(DT_NOTIF, format = "%Y-%m-%d"), unit = AgTimeUnit)
           ) %>%
    ungroup %>%
    select(-DT_NOTIF) %>%
    mutate(GENDER = ifelse(GENDER == "F", "Female", 
                         ifelse(GENDER == "M", "Male", NA))) %>%
    select(DATE, LEVEL, CODE, AGE_CAT, GENDER, Falciparum, Vivax) %>%
    gather(key = 'TYPE', value = 'CASES', -c(DATE, LEVEL, CODE, AGE_CAT, GENDER))
  
                             
  sivep_strat <- rbind(day_malaria_type_brazil_age_gender, day_malaria_type_state_age_gender, day_malaria_type_age_gender)
  
  return(sivep_strat)

}

```






- see pictures to do the figures for the paper
- figure age-gender use the script for pregnancy in PTR proposal- ask Narimane.
- figure por pv/pf- ask Narimane for the script

- For the Pv/Pf figure, you can find it in the "MALARIA_TS_PLOTS" script, lines 1422-1446



```{r}


```

