---
title: "Malaria_Cases_API_final_functions"
author: "Raquel Lana and Narimane Nekkab"
date: "4 de fevereiro de 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Final functions for the analyzes and figures of the first paper

- Packages - move to header file together the parameters.
```{r}
library("ggplot2")
library("forecast")
library("tseries")
library("stringr")
library("gridExtra")
library("data.table")
library("fUnitRoots")
library("FitAR")
library("lubridate")
#library("parallel")
library("TSA")
library("dplyr")
library("foreach")
library("tstools")
library("ggfortify")
library("RColorBrewer")
library("colorspace")
library("foreign")
library("tidyverse")
library("maptools")
library("rgdal")

```

- mover para o header file para rodar a função abaixo
```{r}
# Get SIVEP data cleaned up and by species
SIVEPFilePath = "~/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/SIVEP_clean.RData"
SIVEPFilePath = "~/Documentos/PosDoc_PROCC/OwnCloud/malaria/AnalisesRaquel/Pasteur/SIVEP_clean.RData"
StartYear = 2004 # (inferior limit = 2003)
EndYear = 2017 # (superior limit = 2018)
AgTimeUnit = "week" # ("week", "month")
byNotification = T 
byResidence = F
byInfection = F
byPCD = F
byPCDACD = T
d <- getWEEKLY_SIVEP_MALARIA_TYPE(SIVEPFilePath, StartYear, EndYear, AgTimeUnit)

```

### Function to organize the SIVEP database per:
- DATE: from the date of notification being that it can provide the first day of the epidemiological week in which the notification date was made or the month of this date.    
- LEVEL: Country (BR), State (UF), Municipality (MU).   
- CODE: Country (1), State (Amazon Basin states IBGE codes) and Municipality (Amazon Basin municipalities IBGE codes).  
- TYPE: Falciparum ou Vivax malaria
- Cases: number of cases counted by date, level, code and type of malaria.  
```{r}
getWEEKLY_SIVEP_MALARIA_TYPE <- function(SIVEPFilePath, StartYear, EndYear, AgTimeUnit){
  
  # Get SIVEP raw notification data
  load(SIVEPFilePath)

  # Choose time period
  df <- df[which(df[,"DT_NOTIF"] >= paste0(StartYear,"-01-01") & df[,"DT_NOTIF"] <= paste0(EndYear, "-12-31")),]
  
  
  # Replace missing state code with municipality state code
  df$UF_NOTIF=substr(df$MUN_NOTI, 1, 2) 
  df$UF_RESID=substr(df$MUN_RESI, 1, 2)
  df$UF_INFEC=substr(df$MUN_INFE, 1, 2)

  # Select administrative level via code
  if(byNotification){
    df$PAIS_CODE <- 1
    names(df)[names(df) == 'UF_NOTIF'] <- 'UF_CODE'
    names(df)[names(df) == 'MUN_NOTI'] <- 'MUN_CODE'
  }
  if(byResidence){
    names(df)[names(df) == 'PAIS_RES'] <- 'PAIS_CODE'
    names(df)[names(df) == 'UF_RESID'] <- 'UF_CODE'
    names(df)[names(df) == 'MUN_RESI'] <- 'MUN_CODE'
  }
  if(byInfection){
    names(df)[names(df) == 'PAIS_INF'] <- 'PAIS_CODE'
    names(df)[names(df) == 'UF_INFEC'] <- 'UF_CODE'
    names(df)[names(df) == 'MUN_INFE'] <- 'MUN_CODE'
  }
  
  #Select the cases by Type of Detection
  if(byPCD){
    df <- df %>%
      filter(TIPO_LAM == "PCD")
  }
  if(byPCDACD){
    df <- df %>%
      filter(TIPO_LAM == "PCD" | TIPO_LAM == "ACD")
  }
  
  # Split data by malaria type and aggregate to MU - level
  week_malaria_type_muni <- df %>%
    group_by(DT_NOTIF, MUN_CODE) %>%
    count(RES_EXAM) %>%
    mutate(LEVEL = "MU") %>%
    select(DT_NOTIF, LEVEL, MUN_CODE, RES_EXAM, n) %>%
    spread(RES_EXAM, n, fill = 0) %>%
    rename(CODE = MUN_CODE,
           FALCI = "Falciparum",
           VF = "V+F",
           VIVAX = "Vivax") %>%
    mutate(Falciparum = FALCI + VF,
           Vivax = VIVAX + VF,
           DATE = floor_date(as.Date(DT_NOTIF, format = "%Y-%m-%d"), unit = AgTimeUnit)
           ) %>%
    ungroup %>%
    select(-DT_NOTIF) %>%
    select(DATE, LEVEL, CODE, Falciparum, Vivax) %>%
    gather(key = 'TYPE', value = 'CASES', -c(DATE, LEVEL, CODE))

  
  # UF - level
  week_malaria_type_state <- df %>%
    group_by(DT_NOTIF, UF_CODE) %>%
    count(RES_EXAM) %>%
    mutate(LEVEL = "UF") %>%
    select(DT_NOTIF, LEVEL, UF_CODE, RES_EXAM, n) %>%
    spread(RES_EXAM, n, fill = 0) %>%
    rename(CODE = UF_CODE,
           FALCI = "Falciparum",
           VF = "V+F",
           VIVAX = "Vivax") %>%
    mutate(Falciparum = FALCI + VF,
           Vivax = VIVAX + VF,
           DATE = floor_date(as.Date(DT_NOTIF, format = "%Y-%m-%d"), unit = AgTimeUnit)
           ) %>%      
    ungroup %>%
    select(-DT_NOTIF) %>%
    select(DATE, LEVEL, CODE, Falciparum, Vivax) %>%
    gather(key = 'TYPE', value = 'CASES', -c(DATE, LEVEL, CODE))
  
  # BR - level
  week_malaria_type_brazil <- df %>%
    group_by(DT_NOTIF, PAIS_CODE) %>%
    count(RES_EXAM) %>%
    mutate(LEVEL = "BR")  %>% 
    select(DT_NOTIF, LEVEL, PAIS_CODE, RES_EXAM, n) %>%
    spread(RES_EXAM, n, fill = 0) %>%
    rename(CODE = PAIS_CODE,
           FALCI = "Falciparum",
           VF = "V+F",
           VIVAX = "Vivax") %>%
    mutate(Falciparum = FALCI + VF,
           Vivax = VIVAX + VF,
           DATE = floor_date(as.Date(DT_NOTIF, format = "%Y-%m-%d"), unit = AgTimeUnit)
           ) %>%
    ungroup %>%
    select(-DT_NOTIF) %>%
    select(DATE, LEVEL, CODE, Falciparum, Vivax) %>%
    gather(key = 'TYPE', value = 'CASES', -c(DATE, LEVEL, CODE))
    
  # Merge together
  SIVEP_MALARIA_TYPE <- rbind(week_malaria_type_brazil, week_malaria_type_state, week_malaria_type_muni)
  rm(df, treat, week_malaria_type_brazil, week_malaria_type_state, week_malaria_type_muni)


  return(SIVEP_MALARIA_TYPE)
  
}


```

### Function to agreggate the population size from IBGE and calculate API.
- Move to header file 
```{r}
#IBGEFilePath = "~/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Malaria_Mapping_TimeSeries_Data/BRA_POP_EST.csv"
IBGEFilePath = "~/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Malaria_Mapping_TimeSeries_Data/BRA_POP_EST.csv"
#StartYear = 2004 # (inferior limit = 2003)
#EndYear = 2017 # (superior limit = 2018)
#AgTimeUnit = "week" # ("week", "month")
d2 <- getWEEKLY_SIVEP_API(IBGEFilePath, StartYear, EndYear, AgTimeUnit)

```

- NAME: name of the administrative levels.  
- YEAR: year of the notification and population estimated.  
- POP_SIZE: population estimated by year and administrative levels
- API: annual parasite index - ver se tem que dividir por semanas pois se chama annual p. in.
```{r}
getWEEKLY_SIVEP_API <- function(IBGEFilePath, StartYear, EndYear, AgTimeUnit){
  
  # Get IBGE population size 
  d1 <- read.csv(IBGEFilePath, stringsAsFactors = F)

  # Reshape the population size table to join with cases table
  d1 <- d1 %>%
    select(-GR) %>%
    filter(!LEVEL %in% c("GR")) %>%
    mutate(CODE = replace(CODE, CODE == 0, 1)) %>%
    gather(key = 'YEAR', value = 'POP_SIZE', -c(LEVEL, CODE, NAME)) %>%
    mutate(YEAR = as.integer(str_sub(YEAR, start = 2, end = 5))) %>%
    filter(YEAR >= StartYear & YEAR <= EndYear)
  
  # Create the column YEAR in cases table to merge with population size
  d <- d %>%
    mutate(YEAR = year(DATE),
           CODE = as.numeric(CODE))

  # Join cases and population tables
  API_MALARIA <- merge(d1, d)
    
  # Calculate API
  API_MALARIA <- API_MALARIA %>%
    mutate(API = CASES/POP_SIZE * 1000)
  
  # Reorder and rearrange variables
  API_MALARIA <- API_MALARIA %>%
    select(LEVEL, CODE, NAME, DATE, YEAR, TYPE, CASES, POP_SIZE, API) %>%
    arrange(DATE, YEAR)
  
  return(API_MALARIA)
  
}

```


# Function to calculate Pf/Pv and Pv/Total  ratios
- Check if here It's necessary to group by YEAR.
- Obs.: for some strange reason **tidyr** needs another index to do the **spread**.
```{r}
d3 <- getWEEKLY_SIVEP_RATIOS(API_MALARIA, StartYear, EndYear, AgTimeUnit) # move to header file
  
getWEEKLY_SIVEP_RATIOS <- function(API_MALARIA, StartYear, EndYear, AgTimeUnit){

  # Spreading to facilitate the ratio's calculus
  RATIOS <- d2 %>%
  select(LEVEL, CODE, NAME, DATE, YEAR, TYPE, CASES) %>%
  group_by_at(vars(-CASES)) %>%  # group by everything other than the value column. 
  mutate(row_id = 1:n()) %>% 
  ungroup() %>%  # build group index
  spread(key = TYPE, value = CASES, fill = 0) %>%    # spread
  select(-row_id) %>%  # drop the index
  mutate(PvPf = Falciparum / Vivax,
         PvTOTAL = Vivax / (Falciparum + Vivax)
         ) %>%
  select(LEVEL, CODE, NAME, DATE, YEAR, PvPf, PvTOTAL)  
  
  return(RATIOS)  
  
}

```



### Municipalities plots fro each state
- Ler a tabela de ADMIn anames -facilita pra plotar o gráfico
- Não consegui salvar ainda. Try to save again. Doing
```{r}
ADMIN_NAMES <- read.table("~/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Malaria_Mapping_TimeSeries_Data/BRA_ADMIN_NAMES.csv", sep = " ")
head(ADMIN_NAMES)

ADMIN_NAMES <- ADMIN_NAMES %>%
  ungroup %>%
  filter(Level != "GR") %>%
  mutate(Code = ifelse(Code == 0, 1, Code))

# Assign names
d2$STATE = ADMIN_NAMES[match(d2$CODE, ADMIN_NAMES$Code),"UF"] 
d2$NAME = ADMIN_NAMES[match(d2$CODE, ADMIN_NAMES$Code),"Name"]

# States
States_sivep = c("AC","AM","AP","MA","MT","PA","RO","RR","TO")

MUN_Plots = foreach(i = 1:length(States_sivep)) %do% {
  Plot_Name = paste0("TS_Plot_", States_sivep[i])
  assign(Plot_Name, ggplot(data = subset(d2, LEVEL == "MU" & 
                                           STATE == States_sivep[i]), 
                           aes(DATE, CASES, color = TYPE)) +
           stat_summary(fun.y = sum, geom = "line") +
           scale_x_date(breaks = "year", 
                        date_labels = "%Y") +
           scale_color_manual(values = c("#31a354","#3182bd")) +
           facet_wrap(~NAME) +
           labs(title = paste0("P. vivax and P. falciparum cases in ", States_sivep[i]," state by municipality, ", 
                               StartYear, "-", EndYear), x = "Year", y = "Number of Cases") + 
           guides(color = guide_legend(title = "Plasmodium species")) +
           theme(panel.grid.minor.x = element_blank(),
                 axis.text.x = element_text(angle = 90, hjust = 1)))
}




```


### State plots  - try again - doing
```{r}
UF_Plots = foreach(i = 1:length(States_sivep)) %do% {
  Plot_Name = paste0("TS_Plot_", States_sivep[i])
  assign(Plot_Name, ggplot(data = subset(d2, LEVEL == "UF" & STATE == States_sivep[i]), 
                           aes(DATE, CASES, color = TYPE)) +
           stat_summary(fun.y = sum, geom = "line") +
           scale_x_date(breaks = "year", 
                        date_labels = "%Y") +
           scale_color_manual(values = c("#31a354","#3182bd")) +
           facet_wrap(~NAME) +
           labs(title = paste0("P. vivax and P. falciparum cases in ", States_sivep[i], "Brazil", 
                               StartYear, "-", EndYear), x = "Year", y = "Number of Cases") + 
           guides(color=guide_legend(title = "Plasmodium species")) +
           theme(panel.grid.minor.x = element_blank(),
                 axis.text.x = element_text(angle = 90, hjust = 1)))
}

```


- Não consegui plotar API ou casos somado por estado.
- Não consegui salvar as figuras.


## Function to organize the cases per age and gender - doing
- um painel com todos os estados- somatório de todos os anos.
- Para suplementar, seria interessante ter uma figura para cada estado mostrando por ano- facet_wrap por ano.

-REVER SE FAZ POR ANO- VER O PLOT.
```{r}
load("/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/SIVEP_clean.RData") # move to header file

getSTRAT_AGE_GENDER <- function(SIVEPFilePath, StartYear, EndYear, AgTimeUnit){

  
  # Brazil  
  day_malaria_type_brazil_age_gender <- df %>%
    group_by(DT_NOTIF, AGE_CAT, SEXO) %>%
    count(RES_EXAM)%>%
    mutate(LEVEL = "BR",
           CODE = "1") %>%
    select(DT_NOTIF, LEVEL, CODE, RES_EXAM, AGE_CAT, SEXO, n) %>%
    spread(RES_EXAM, n, fill = 0) %>%
    rename(GENDER = SEXO, FALCI = "Falciparum", VF = "V+F", VIVAX = "Vivax") %>%
    mutate(Falciparum = FALCI + VF,
           Vivax = VIVAX + VF,
           DATE = floor_date(as.Date(DT_NOTIF, format = "%Y-%m-%d"), unit = AgTimeUnit)
            ) %>%
    ungroup %>%
    select(-DT_NOTIF) %>%
    mutate(GENDER = ifelse(GENDER == "F", "Female", 
                           ifelse(GENDER == "M", "Male", NA))) %>%
    select(DATE, LEVEL, CODE, AGE_CAT, GENDER, Falciparum, Vivax) %>%
    gather(key = 'TYPE', value = 'CASES', -c(DATE, LEVEL, CODE, AGE_CAT, GENDER))
  
  # State
  day_malaria_type_state_age_gender <- df %>%
    group_by(DT_NOTIF, UF_NOTIF, AGE_CAT, SEXO) %>%
    count(RES_EXAM) %>%
    mutate(LEVEL = "UF") %>%
    select(DT_NOTIF, LEVEL, UF_NOTIF, RES_EXAM, AGE_CAT, SEXO, n) %>%
    spread(RES_EXAM, n, fill = 0) %>%
    rename(CODE = UF_NOTIF, GENDER = SEXO, FALCI = "Falciparum", VF = "V+F", VIVAX = "Vivax") %>%
    mutate(Falciparum = FALCI + VF,
           Vivax = VIVAX + VF,
           DATE = floor_date(as.Date(DT_NOTIF, format = "%Y-%m-%d"), unit = AgTimeUnit)
            ) %>%
    ungroup %>%
    select(-DT_NOTIF) %>%
    mutate(GENDER = ifelse(GENDER == "F", "Female", 
                           ifelse(GENDER == "M", "Male", NA))) %>%
    select(DATE, LEVEL, CODE, AGE_CAT, GENDER, Falciparum, Vivax) %>%
    gather(key = 'TYPE', value = 'CASES', -c(DATE, LEVEL, CODE, AGE_CAT, GENDER))
  
  # Municipality
  day_malaria_type_age_gender <- df %>%
    group_by(DT_NOTIF, MUN_NOTI, AGE_CAT, SEXO) %>%
    count(RES_EXAM) %>%
    mutate(LEVEL = "MU") %>%
    select(DT_NOTIF, LEVEL, MUN_NOTI, RES_EXAM, AGE_CAT, SEXO, n) %>%
    spread(RES_EXAM, n, fill = 0) %>%
    rename(CODE = MUN_NOTI, GENDER = SEXO, FALCI = "Falciparum", VF = "V+F", VIVAX = "Vivax") %>%
    mutate(Falciparum = FALCI + VF,
           Vivax = VIVAX + VF,
           DATE = floor_date(as.Date(DT_NOTIF, format = "%Y-%m-%d"), unit = AgTimeUnit)
           ) %>%
    ungroup %>%
    select(-DT_NOTIF) %>%
    mutate(GENDER = ifelse(GENDER == "F", "Female", 
                         ifelse(GENDER == "M", "Male", NA))) %>%
    select(DATE, LEVEL, CODE, AGE_CAT, GENDER, Falciparum, Vivax) %>%
    gather(key = 'TYPE', value = 'CASES', -c(DATE, LEVEL, CODE, AGE_CAT, GENDER))
  
                             
  sivep_strat <- rbind(day_malaria_type_brazil_age_gender, day_malaria_type_state_age_gender, day_malaria_type_age_gender)
  
  return(sivep_strat)

}

```






- see pictures to do the figures for the paper
- figure age-gender use the script for pregnancy in PTR proposal- ask Narimane.
- figure por pv/pf- ask Narimane for the script

- For the Pv/Pf figure, you can find it in the "MALARIA_TS_PLOTS" script, lines 1422-1446



```{r}


```

