---
title: "Manipulate_data_maps_ts"
author: "Raquel Lana and Narimane"
date: "6 de dezembro de 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Aggregating SIVEP data for year to build API maps

- Packages
```{r}
library(foreign)
library('stringr')
library(tidyverse)
library('data.table')
library('forcats')
library(lubridate)

```

### Loading SIVEP RData
Environment: NN or RL - RL is **F**.
```{r}
envNN = F

if(envNN){
  load("/SIVEP_clean.RData")
}else{load("/home/claudia/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/SIVEP_clean.RData") #Fiocruz desktop
#}else{load("/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/SIVEP_clean.RData") #my pc
  
}

```

## Counting malaria type per date of notification, levels of administrate units (Country, State and Municipality).

### Counting malaria cases by day, UF, municipality and malaria type.
```{r}
day_malaria_type <- df %>%
                      group_by(DT_NOTIF, MUN_NOTI) %>%
                      count(RES_EXAM) %>%
                      mutate(LEVEL = "MU") %>%
                      select(DT_NOTIF, LEVEL, MUN_NOTI, RES_EXAM, n) %>%
                      spread(RES_EXAM, n, fill = 0) %>%
                      rename(CODE = MUN_NOTI) %>%
                      rename(FALCI = "Falciparum") %>%
                      rename(VF = "V+F") %>%
                      rename(VIVAX = "Vivax") %>%
                      mutate(Falciparum = FALCI + VF) %>%
                      mutate(Vivax = VIVAX + VF) %>%
                      select(DT_NOTIF, LEVEL, CODE, Falciparum, Vivax) %>%
                      gather(key = 'TYPE', value = 'CASES', -c(DT_NOTIF, LEVEL, CODE))


```

### Counting malaria cases by day, state and malaria type
```{r}
day_malaria_type_state <- df %>%
                          group_by(DT_NOTIF, UF_NOTIF) %>%
                          count(RES_EXAM) %>%
                          mutate(LEVEL = "UF") %>%
                          select(DT_NOTIF, LEVEL, UF_NOTIF, RES_EXAM, n) %>%
                          spread(RES_EXAM, n, fill = 0) %>%
                          rename(CODE = UF_NOTIF) %>%
                          rename(FALCI = "Falciparum") %>%
                          rename(VF = "V+F") %>%
                          rename(VIVAX = "Vivax") %>%
                          mutate(Falciparum = FALCI + VF) %>%
                          mutate(Vivax = VIVAX + VF) %>%
                          select(DT_NOTIF, LEVEL, CODE, Falciparum, Vivax) %>%
                          gather(key = 'TYPE', value = 'CASES', -c(DT_NOTIF, LEVEL, CODE))

```


### Counting malaria cases by day, country and malaria type
```{r}
day_malaria_type_brazil <- df %>%
                          group_by(DT_NOTIF) %>%
                          count(RES_EXAM)%>%
                          mutate(LEVEL = "BR",
                                 CODE = "1") %>%
                          select(DT_NOTIF, LEVEL, CODE, RES_EXAM, n) %>%
                          spread(RES_EXAM, n, fill = 0) %>%
                          rename(FALCI = "Falciparum") %>%
                          rename(VF = "V+F") %>%
                          rename(VIVAX = "Vivax") %>%
                          mutate(Falciparum = FALCI + VF) %>%
                          mutate(Vivax = VIVAX + VF) %>%
                          select(DT_NOTIF, LEVEL, CODE, Falciparum, Vivax) %>%
                          gather(key = 'TYPE', value = 'CASES', -c(DT_NOTIF, LEVEL, CODE))

```

## Adding GENDER

### Counting malaria cases by day, municipality, GENDER, malaria type.
```{r}
day_malaria_type_gender <- df %>%
                            group_by(DT_NOTIF, MUN_NOTI, SEXO) %>%
                            count(RES_EXAM) %>%
                            mutate(LEVEL = "MU") %>%
                            select(DT_NOTIF, LEVEL, MUN_NOTI, RES_EXAM, SEXO, n) %>%
                            spread(RES_EXAM, n, fill = 0) %>%
                            rename(CODE = MUN_NOTI, FALCI = "Falciparum", VF = "V+F", VIVAX = "Vivax", GENDER = SEXO) %>%
                            mutate(Falciparum = FALCI + VF) %>%
                            mutate(Vivax = VIVAX + VF) %>%
                            select(DT_NOTIF, LEVEL, CODE, GENDER, Falciparum, Vivax) %>%
                            gather(key = 'TYPE', value = 'CASES', -c(DT_NOTIF, LEVEL, CODE, GENDER))


```

### Counting malaria cases by day, state, GENDER and malaria type
```{r}
day_malaria_type_state_gender <- df %>%
                          group_by(DT_NOTIF, UF_NOTIF, SEXO) %>%
                          count(RES_EXAM) %>%
                          mutate(LEVEL = "UF") %>%
                          select(DT_NOTIF, LEVEL, UF_NOTIF, RES_EXAM, SEXO, n) %>%
                          spread(RES_EXAM, n, fill = 0) %>%
                          rename(CODE = UF_NOTIF, FALCI = "Falciparum", VF = "V+F", VIVAX = "Vivax", GENDER = SEXO) %>%
                          mutate(Falciparum = FALCI + VF) %>%
                          mutate(Vivax = VIVAX + VF) %>%
                          select(DT_NOTIF, LEVEL, CODE, GENDER, Falciparum, Vivax) %>%
                          gather(key = 'TYPE', value = 'CASES', -c(DT_NOTIF, LEVEL, CODE, GENDER))

```


### Counting malaria cases by day, country, GENDER and malaria type
```{r}
day_malaria_type_brazil_gender <- df %>%
                          group_by(DT_NOTIF, SEXO) %>%
                          count(RES_EXAM)%>%
                          mutate(LEVEL = "BR",
                                 CODE = "1") %>%
                          select(DT_NOTIF, LEVEL, CODE, RES_EXAM, SEXO, n) %>%
                          spread(RES_EXAM, n, fill = 0) %>%
                          rename(FALCI = "Falciparum", VF = "V+F", VIVAX = "Vivax", GENDER = SEXO) %>%
                          mutate(Falciparum = FALCI + VF) %>%
                          mutate(Vivax = VIVAX + VF) %>%
                          select(DT_NOTIF, LEVEL, CODE, GENDER, Falciparum, Vivax) %>%
                          gather(key = 'TYPE', value = 'CASES', -c(DT_NOTIF, LEVEL, CODE, GENDER))

```


### Counting malaria cases by day, UF, municipality, AGE Categories, malaria type.
```{r}
day_malaria_type_age <- df %>%envNN = F
                            group_by(DT_NOTIF, MUN_NOTI, AGE_CAT) %>%
                            count(RES_EXAM) %>%
                            mutate(LEVEL = "MU") %>%
                            select(DT_NOTIF, LEVEL, MUN_NOTI, RES_EXAM, AGE_CAT, n) %>%
                            spread(RES_EXAM, n, fill = 0) %>%
                            rename(CODE = MUN_NOTI, FALCI = "Falciparum", VF = "V+F", VIVAX = "Vivax") %>%
                            mutate(Falciparum = FALCI + VF) %>%
                            mutate(Vivax = VIVAX + VF) %>%
                            select(DT_NOTIF, LEVEL, CODE, AGE_CAT, Falciparum, Vivax) %>%
                            gather(key = 'TYPE', value = 'CASES', -c(DT_NOTIF, LEVEL, CODE, AGE_CAT))


```

### Counting malaria cases by day, UF, AGE RANGE and malaria type
```{r}
day_malaria_type_state_age <- df %>%
                          group_by(DT_NOTIF, UF_NOTIF, AGE_CAT) %>%
                          count(RES_EXAM) %>%
                          mutate(LEVEL = "UF") %>%
                          select(DT_NOTIF, LEVEL, UF_NOTIF, RES_EXAM, AGE_CAT, n) %>%
                          spread(RES_EXAM, n, fill = 0) %>%
                          rename(CODE = UF_NOTIF, FALCI = "Falciparum", VF = "V+F", VIVAX = "Vivax") %>%
                          mutate(Falciparum = FALCI + VF) %>%
                          mutate(Vivax = VIVAX + VF) %>%
                          select(DT_NOTIF, LEVEL, CODE, AGE_CAT, Falciparum, Vivax) %>%
                          gather(key = 'TYPE', value = 'CASES', -c(DT_NOTIF, LEVEL, CODE, AGE_CAT))

```


### Counting malaria cases by day, AGE categories and malaria type
```{r}
day_malaria_type_brazil_age <- df %>%
                          group_by(DT_NOTIF, AGE_CAT) %>%
                          count(RES_EXAM)%>%
                          mutate(LEVEL = "BR",
                                 CODE = "1") %>%
                          select(DT_NOTIF, LEVEL, CODE, RES_EXAM, AGE_CAT, n) %>%
                          spread(RES_EXAM, n, fill = 0) %>%
                          rename(FALCI = "Falciparum", VF = "V+F", VIVAX = "Vivax") %>%
                          mutate(Falciparum = FALCI + VF) %>%
                          mutate(Vivax = VIVAX + VF) %>%
                          select(DT_NOTIF, LEVEL, CODE, AGE_CAT, Falciparum, Vivax) %>%
                          gather(key = 'TYPE', value = 'CASES', -c(DT_NOTIF, LEVEL, CODE, AGE_CAT))

```

### Checking registers per notification, residence and probable infection place.
- UF_NOTIF: 9 
- UF_RESID: 27 + missings
- UF_INFEC: 27 + missings
```{r}
length(unique(df$UF_NOTIF))
length(unique(df$UF_RESID))
length(unique(df$UF_INFEC))

```

- Different visualizations per state. 
```{r}
NOTIF_RESID <- df %>% 
                group_by(UF_NOTIF, UF_RESID) %>%
                count(RES_EXAM)# %>%
#                filter(is.na(UF_RESID))


NOTIF_INFEC<- df %>% 
                group_by(UF_NOTIF, UF_INFEC) %>%
                count(RES_EXAM)# %>%
#                filter(is.na(UF_INFEC))

RESID_INFEC <- df %>% 
                group_by(UF_RESID, UF_INFEC) %>%
                count(RES_EXAM)# %>%
#                filter(is.na(UF_INFEC))
  
ALL_UF <- df %>% 
            group_by(UF_NOTIF, UF_RESID, UF_INFEC) %>%
            count(RES_EXAM)

```

## Malaria per notification, residence and probable infection places

### Counting malaria cases per day, notification, residence, probable infection per country and malaria type.
```{r}
day_malaria_type_places_country <- df %>%
                          group_by(DT_NOTIF, PAIS_RES, PAIS_INF) %>%
                          count(RES_EXAM) %>%
                          mutate(LEVEL = "BR",
                                 PAIS_NOTI = "1") %>%
                          select(DT_NOTIF, LEVEL, PAIS_NOTI, PAIS_RES, PAIS_INF, RES_EXAM, n) %>%
                          spread(RES_EXAM, n, fill = 0) %>%
                          rename(CODE_NOTIF = PAIS_NOTI, CODE_RESID = PAIS_RES, CODE_INFEC = PAIS_INF, FALCI = "Falciparum", VF = "V+F", VIVAX = "Vivax") %>%
                          mutate(Falciparum = FALCI + VF) %>%
                          mutate(Vivax = VIVAX + VF) %>%
                          select(DT_NOTIF, LEVEL, CODE_NOTIF, CODE_RESID, CODE_INFEC, Falciparum, Vivax) %>%
                          gather(key = 'TYPE', value = 'CASES', -c(DT_NOTIF, LEVEL, CODE_NOTIF, CODE_RESID, CODE_INFEC))

```


### Counting malaria cases per day, notification, residence, probable infection per state and malaria type.
```{r}
day_malaria_type_places_state <- df %>%
                          group_by(DT_NOTIF, UFenvNN = F_NOTIF, UF_RESID, UF_INFEC) %>%
                          count(RES_EXAM) %>%
                          mutate(LEVEL = "UF") %>%
                          select(DT_NOTIF, LEVEL, UF_NOTIF, UF_RESID, UF_INFEC, RES_EXAM, n) %>%
                          spread(RES_EXAM, n, fill = 0) %>%
                          rename(CODE_NOTIF = UF_NOTIF, CODE_RESID = UF_RESID, CODE_INFEC = UF_INFEC, FALCI = "Falciparum", VF = "V+F", VIVAX = "Vivax") %>%
                          mutate(Falciparum = FALCI + VF) %>%
                          mutate(Vivax = VIVAX + VF) %>%
                          select(DT_NOTIF, LEVEL, CODE_NOTIF, CODE_RESID, CODE_INFEC, Falciparum, Vivax) %>%
                          gather(key = 'TYPE', value = 'CASES', -c(DT_NOTIF, LEVEL, CODE_NOTIF, CODE_RESID, CODE_INFEC))


```


### Counting malaria cases per day, notification, residence, probable infection per municipality and malaria type.
```{r}
day_malaria_type_places_muni <- df %>%
                          group_by(DT_NOTIF, MUN_NOTI, MUN_RESI, MUN_INFE) %>%
                          count(RES_EXAM) %>%
                          mutate(LEVEL = "MU") %>%
                          select(DT_NOTIF, LEVEL, MUN_NOTI, MUN_RESI, MUN_INFE, RES_EXAM, n) %>%
                          spread(RES_EXAM, n, fill = 0) %>%
                          rename(CODE_NOTIF = MUN_NOTI, CODE_RESID = MUN_RESI, CODE_INFEC = MUN_INFE, FALCI = "Falciparum", VF = "V+F", VIVAX = "Vivax") %>%
                          mutate(Falciparum = FALCI + VF) %>%
                          mutate(Vivax = VIVAX + VF) %>%
                          select(DT_NOTIF, LEVEL, CODE_NOTIF, CODE_RESID, CODE_INFEC, Falciparum, Vivax) %>%
                          gather(key = 'TYPE', value = 'CASES', -c(DT_NOTIF, LEVEL, CODE_NOTIF, CODE_RESID, CODE_INFEC))


```


# Case per age and gender population size - API

### Counting malaria cases by day, municipality, age categories, gender, malaria type.
```{r}
day_malaria_type_age_gender <- df %>%
                            group_by(DT_NOTIF, MUN_NOTI, AGE_CAT, SEXO) %>%
                            count(RES_EXAM) %>%
                            mutate(LEVEL = "MU") %>%
                            select(DT_NOTIF, LEVEL, MUN_NOTI, RES_EXAM, AGE_CAT, SEXO, n) %>%
                            spread(RES_EXAM, n, fill = 0) %>%
                            rename(CODE = MUN_NOTI, GENDER = SEXO, FALCI = "Falciparum", VF = "V+F", VIVAX = "Vivax") %>%
                            mutate(Falciparum = FALCI + VF) %>%
                            mutate(Vivax = VIVAX + VF) %>%
                            select(DT_NOTIF, LEVEL, CODE, AGE_CAT, GENDER, Falciparum, Vivax) %>%
                            gather(key = 'TYPE', value = 'CASES', -c(DT_NOTIF, LEVEL, CODE, AGE_CAT, GENDER))
                           

```

### Counting malaria cases by day, state, age caterories, gender and malaria type
```{r}
day_malaria_type_state_age_gender <- df %>%
                          group_by(DT_NOTIF, UF_NOTIF, AGE_CAT, SEXO) %>%
                          count(RES_EXAM) %>%
                          mutate(LEVEL = "UF") %>%
                          select(DT_NOTIF, LEVEL, UF_NOTIF, RES_EXAM, AGE_CAT, SEXO, n) %>%
                          spread(RES_EXAM, n, fill = 0) %>%
                          rename(CODE = UF_NOTIF, GENDER = SEXO, FALCI = "Falciparum", VF = "V+F", VIVAX = "Vivax") %>%
                          mutate(Falciparum = FALCI + VF) %>%
                          mutate(Vivax = VIVAX + VF) %>%
                          select(DT_NOTIF, LEVEL, CODE, AGE_CAT, GENDER, Falciparum, Vivax) %>%
                          gather(key = 'TYPE', value = 'CASES', -c(DT_NOTIF, LEVEL, CODE, AGE_CAT, GENDER))

```


### Counting malaria cases by day, country, age categories, gender and malaria type
```{r}
day_malaria_type_brazil_age_gender <- df %>%
                          group_by(DT_NOTIF, AGE_CAT, SEXO) %>%
                          count(RES_EXAM)%>%
                          mutate(LEVEL = "BR",
                                 CODE = "1") %>%
                          select(DT_NOTIF, LEVEL, CODE, RES_EXAM, AGE_CAT, SEXO, n) %>%
                          spread(RES_EXAM, n, fill = 0) %>%
                          rename(GENDER = SEXO, FALCI = "Falciparum", VF = "V+F", VIVAX = "Vivax") %>%
                          mutate(Falciparum = FALCI + VF) %>%
                          mutate(Vivax = VIVAX + VF) %>%
                          select(DT_NOTIF, LEVEL, CODE, AGE_CAT, GENDER, Falciparum, Vivax) %>%
                          gather(key = 'TYPE', value = 'CASES', -c(DT_NOTIF, LEVEL, CODE, AGE_CAT, GENDER))

```

# rbind population table per level
- Rename GENDER and exclude "I" and NA's
```{r}
sivep_strat <- rbind(day_malaria_type_brazil_age_gender, day_malaria_type_state_age_gender, day_malaria_type_age_gender)

sivep_strat$GENDER1 <- NA
sivep_strat[which(sivep_strat$GENDER == "F"),]$GENDER1 <- "Female" 
sivep_strat[which(sivep_strat$GENDER == "M"),]$GENDER1 <- "Male" 
sivep_strat[which(sivep_strat$GENDER == "I"),]$GENDER1 <- NA 

unique(sivep_strat$GENDER1)
class(sivep_strat$GENDER1)
table(sivep_strat$GENDER1)

sivep_strat <- sivep_strat[, c(1, 2, 3, 4, 8, 6, 7)]

sivep_strat <- sivep_strat %>%
                        rename(GENDER = GENDER1)

write.csv(sivep_strat, file = "/home/claudia/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Malaria_Mapping_TimeSeries_Data/sivep_admlevels_dtnotif_STRAT.csv", row.names = F)
```



# Population size per level, code, age categories and gender
```{r}
if(envNN){
  pop_strat <- read.csv("/BRA_AGE_SEX_MUNICIP_2010.csv")
}else{pop_strat <- fread("/home/claudia/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Malaria_Mapping_TimeSeries_Data/BRA_AGE_SEX_MUNICIP_2010.csv", stringsAsFactors = F, encoding = 'Latin-1')}

#pop_strat <- fread("/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Malaria_Mapping_TimeSeries_Data/BRA_AGE_SEX_MUNICIP_2010.csv", stringsAsFactors = F, encoding = 'Latin-1') #my pc

pop_strat <- pop_strat[, -5]
pop_strat 

head(pop_strat)
unique(pop_strat$Level) 
unique(pop_strat$Code) 
unique(pop_strat$Group) 
summary(pop_strat)


```

- Recategorizing age categories and cleaning
```{r}
pop_strat <- pop_strat %>%
                  filter(Group == "0 a 4 anos" | Group == "5 a 9 anos" | Group == "10 a 14 anos" | Group == "15 a 19 anos" | Group == "20 a 24 anos" | Group == "25 a 29 anos" | Group == "30 a 34 anos" | Group == "35 a 39 anos" | Group == "40 a 44 anos" | Group == "45 a 49 anos"  | Group == "50 a 54 anos" | Group == "55 a 59 anos" | Group == "60 a 64 anos" | Group == "65 a 69 anos" | Group == "70 a 74 anos" | Group == "75 a 79 anos" | Group == "80 a 84 anos" | Group == "85 a 89 anos" | Group == "90 a 94 anos" | Group == "95 a 99 anos" | Group == "100 anos ou mais") %>%
                  mutate(Group = ifelse(Group == "80 anos ou mais" | Group == "80 a 84 anos" | Group == "85 a 89 anos" | Group == "90 a 94 anos" | Group == "95 a 99 anos" | Group == "100 anos ou mais", "80 anos ou mais", Group))


# Sum lines with the category "80 anos ou mais"   
pop_strat$Men <- as.integer(pop_strat$Men)
pop_strat$Women <- as.integer(pop_strat$Women)
pop_strat <- aggregate(cbind(Men, Women) ~ Level + Code + Name + Group, data = pop_strat, FUN = sum)

pop_strat %>% filter(Code == "4304101")  
 
pop_strat <- pop_strat %>%
                   mutate(Group = factor(Group, levels= c("0 a 4 anos", "5 a 9 anos", "10 a 14 anos", "15 a 19 anos", "20 a 24 anos", "25 a 29 anos", "30 a 34 anos",  "35 a 39 anos", "40 a 44 anos", "45 a 49 anos", "50 a 54 anos", "55 a 59 anos", "60 a 64 anos", "65 a 69 anos", "70 a 74 anos", "75 a 79 anos", "80 anos ou mais"), label=c("0-4", "5-9", "10-14", "15-19", "20-24", "25-29", "30-34", "35-39", "40-44", "45-49", "50-54", "55-59", "60-64", "65-69", "70-74", "75-79", "80+"))) %>%
                  filter(Level != "GR") %>%
                  rename(LEVEL = Level, CODE = Code, NAME = Name, AGE_CAT = Group, Male = Men, Female = Women) %>%
                  gather(key = 'GENDER', value = 'POP_SIZE', -c(LEVEL, CODE, NAME, AGE_CAT)) %>%
                  mutate(CODE = str_sub(CODE, 1, 6))

pop_strat %>% filter(CODE == "310010")
pop_strat[which(pop_strat$CODE == "310010" & is.na(pop_strat$AGE_CAT)),]

write.csv(pop_strat, file = "pop_strat_2010.csv", row.names = F)
  
```

Proportion of males and females by gender
```{r}
pop_strat <- read.csv("/home/claudia/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Malaria_Mapping_TimeSeries_Data/pop_strat_2010.csv", stringsAsFactors = F)
head(pop_strat)

pop_strat <- pop_strat %>% 
                    add_count(CODE, wt = POP_SIZE) %>%
                    rename(POP_TOTAL = n) %>%
                    mutate(PROPORTION = POP_SIZE/POP_TOTAL) %>%
                    mutate(PROPORTION = as.numeric(PROPORTION))

pop_strat %>% filter(CODE == "310010")

write.csv(pop_strat, file = "/home/claudia/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Malaria_Mapping_TimeSeries_Data/pop_strat_2010_prop.csv", row.names = F)

```

- Importing pop size table to apply the pop proportion by age and gender per admin. strat and year
```{r}
pop_strat <- read.csv("/home/claudia/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Malaria_Mapping_TimeSeries_Data/pop_strat_2010_prop.csv")
head(pop_strat)

pop_strat %>% filter(POP_TOTAL == 8366)

```

- Organizing the pop table per year for the format LEVEL/CODE/NAME
```{r}
pop_ym <- read.csv("/home/claudia/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Malaria_Mapping_TimeSeries_Data/pop_estimated.csv")
head(pop_ym)


```

### Population size per Level, Code, Name and Year: Municipality
- Obs.: I didn't remane NAME1 to NAME because the difference between the pattern in the other tables.
```{r}
pop_Municipality <- pop_ym %>%
                mutate(LEVEL = "MU") %>%
                rename(CODE = COD.MUNIC1, NAME = NOME.MUNIC) %>%
                select(LEVEL, CODE, NAME, POP.EST, YEAR)

head(pop_Municipality)              
unique(pop_Municipality$LEVEL)

```

### Population size per Level, Code, Name and Year: State
```{r}
pop_State <- pop_ym %>%
                group_by(COD.UF, UF, YEAR) %>%
                summarise(POP.EST = sum(POP.EST)) %>%
                mutate(LEVEL = "UF") %>%
                rename(CODE = COD.UF, NAME = UF) %>%
                select(LEVEL, CODE, NAME, POP.EST, YEAR)

```


### Population size per Level, Code, Name and Year: Brazil
```{r}
pop_Brazil <- pop_ym %>%
                group_by(YEAR) %>%
                summarise(POP.EST = sum(POP.EST)) %>%
                mutate(LEVEL = "BR",
                       CODE = 1,
                       NAME = "Brazil") %>%
                mutate(CODE = as.integer(CODE),
                       NAME = as.factor(NAME)) %>%
                select(LEVEL, CODE, NAME, POP.EST, YEAR)

```

- cbind pop size tables
```{r}
names(pop_Brazil)
names(pop_State)
names(pop_Municipality)
pop_Brazil <- as.data.frame(pop_Brazil)
pop_State <- as.data.frame(pop_State)
pop_Municipality <- as.data.frame(pop_Municipality)

str(pop_Brazil)
str(pop_State)
str(pop_Municipality)

pop_size <- data.frame(rbind(pop_Brazil, pop_State, pop_Municipality))
head(pop_size)

write.csv(pop_size, file = "/home/claudia/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Malaria_Mapping_TimeSeries_Data/pop_size_adm_levels_years.csv", row.names = F)

```


- Join pop size and pop strat
```{r}
dim(pop_strat)
str(pop_strat)
dim(pop_size)
str(pop_size)
pop_strat <- pop_strat[,-3]

pop_size_strat <- left_join(x = pop_strat, y = pop_size)
head(pop_size_strat)
dim(pop_size_strat)

```  

### Calculating pop size per gender, age, adm level and year
```{r}
pop_size_strat <- pop_size_strat %>%
                              mutate(POP.EST.Y = PROPORTION*POP.EST) %>%
                              mutate(POP.EST.Y = round(POP.EST.Y, 0),
                                     LEVEL = factor(LEVEL, c("BR", "UF", "MU"), c("BR", "UF", "MU"))) %>%
                              arrange(CODE) %>%
                              select(LEVEL, CODE, AGE_CAT, GENDER, NAME, YEAR, POP.EST.Y)
                              

# I could not save with this category order.
#pop_size_strat1$AGE_CAT %>% fct_relevel("0-4", "5-9", "10-14", "15-19", "20-24", "25-29", "30-34", "35-39", "40-44", "45-49", "50-54", "55-59", "60-64", "65-69", "70-74", "75-79", "80+" ) 

head(pop_size_strat)
tail(pop_size_strat)


write.csv(pop_size_strat, file = "/home/claudia/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Malaria_Mapping_TimeSeries_Data/pop_size_admlevels_year_STRAT.csv", row.names = F)

```

### Join pop size by adm levels, year and strat. with SIVEP.
- I saved the tables, clear the environment and imported again not to be so slow.
- I have to create the variable year in sivep table, to link with pop table. Faster that run again the tables including year and week of the original database.
- Problem: GENDER in SIVEP has NA's - I removed these lines from the database such as YEAR == 2018. 
- Some municipalities don't have all AGE_CAT, but I couldn't get to visualize it.
- The ideal table is a table with all AGE_CAT and GENDER in all weeks and year, but I still couldn't get to create it. 
```{r}
pop_strat <- read.csv("/home/claudia/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Malaria_Mapping_TimeSeries_Data/pop_size_admlevels_year_STRAT.csv")
sivep_strat <- read.csv("/home/claudia/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Malaria_Mapping_TimeSeries_Data/sivep_admlevels_dtnotif_STRAT.csv")

length(unique(sivep_strat$CODE[sivep_strat$LEVEL == "MU"]))
#712
length(unique(pop_strat$CODE[pop_strat$LEVEL == "MU"]))
#5565

sivep_strat <- sivep_strat %>%
                          mutate(WEEK = epiweek(DT_NOTIF),
                                 YEAR = as.integer(str_sub(DT_NOTIF, start = 1, end = 4))
                                 )

length(sivep_strat$YEAR == 2018)
length(sivep_pop_strat$YEAR == 2018)

sivep_strat1 <- sivep_strat %>%
                            drop_na(GENDER) %>%
                            filter(YEAR <= 2017) %>%
                            group_by(LEVEL, CODE, AGE_CAT, GENDER, TYPE, WEEK, YEAR) %>%
                            summarise(CASES = sum(CASES)) %>%
                            arrange(CODE, AGE_CAT, WEEK, YEAR)
                            
#teste <- sivep_strat1 %>% group_by(CODE) %>% count(AGE_CAT) %>% filter(CODE == 510895)


sivep_pop_strat <- left_join(x = sivep_strat1, y = pop_strat)
sivep_pop_strat <- sivep_pop_strat %>%
                                  mutate(API = CASES/POP.EST.Y * 1000)
                                  

# sivep_strat[which(sivep_strat$AGE_CAT == "0-4" & sivep_strat$CASES == 0),]
# sivep_strat[which(sivep_strat$AGE_CAT == "0-4" & is.na(sivep_strat$CASES)),]
# 
# pop_strat[which(pop_strat$POP.EST.Y == 0),]
# pop_strat[which(pop_strat$AGE_CAT == "0-4" & is.na(pop_strat$POP.EST.Y)),]
# 
# head(sivep_pop_strat1[which(sivep_pop_strat1$AGE_CAT == "0-4" & sivep_pop_strat1$CASES == 0),])
# sivep_strat[which(sivep_strat$AGE_CAT == "0-4" & is.na(sivep_strat$CASES)),]
# 
# head(sivep_pop_strat2[which(is.na(sivep_pop_strat2$POP.EST.Y)),])


```

testing relevel AGE_CAT
```{r}
cat_order <- c("0-4", "5-9", "10-14", "15-19", "20-24", "25-29", "30-34", "35-39", "40-44", "45-49", "50-54", "55-59", "60-64", "65-69", "70-74", "75-79", "80+")

sivep_pop_strat <- sivep_pop_strat %>% 
                        mutate(AGE_CAT1 = factor(AGE_CAT, levels = cat_order)) 

sivep_pop_strat <- sivep_pop_strat[, c(1, 2, 12, 4:11)]
sivep_pop_strat1 <- sivep_pop_strat %>%
                                   rename(AGE_CAT = AGE_CAT1)

write.csv(sivep_pop_strat1, file = "/home/claudia/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Malaria_Mapping_TimeSeries_Data/API_admlevels_year_STRAT.csv", row.names = F)

AG <- ggplot(aes(y = API, x = AGE_CAT, fill = GENDER), data = sivep_pop_strat1[sivep_pop_strat1$CODE == 1,]) +
            geom_bar(stat = "identity", position = "dodge") +
            facet_wrap(~YEAR) +
            theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

```



- abrir o shape de assentamentos e ver se tem de grandes empreendimentos e suas infos.
