---
title: "Initial descriptions paper"
author: "Raquel Lana"
date: "April 17, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Descriptions for Results part 1: manuscript and table 1
### The first initial descriptions for Results section part 1 needs the full database to calculate cases and API for the manuscript and table 1.

- Loading files and libraries
```{r}
library(foreign)
library('stringr')
library("tidyverse")
library(lubridate)
library(data.table)
library(bit64)
require(maptools, quietly = TRUE, warn.conflicts = FALSE)
library('maptools')
library(rgdal)

#load("~/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/RData_full.RData") #until first weeks of 2018
load("/home/rlana/Dados/RData_full_new.RData") #until the end of 2018 and a update of 2017.
#load("/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/AnaliseExploratoria/RData_full_new.RData")

pop <-fread("/home/rlana/Dados/BRA_POP_EST.csv", encoding = "UTF-8")
#pop <-fread("/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/AnaliseExploratoria/BRA_POP_EST.csv", encoding = "UTF-8")

```

## For paper description  

### Brazil 2003 to 2018:  
```{r}
tmal <- df %>%
  filter(ANO != 2003) %>% #Comment this line if keep 2003
  filter(RES_EXAM != "0") %>%  #Até aqui são: 5496673 casos 
  mutate(RES_EXAM = if_else(RES_EXAM == "2" | RES_EXAM == "3" | RES_EXAM == "7", "F", 
          if_else(RES_EXAM == "4" | RES_EXAM == "6", "V",
                  if_else(RES_EXAM == "5", "F+V",
                          if_else(RES_EXAM == "8", "M", 
                                  if_else(RES_EXAM == "9", "F+M",
                                          if_else(RES_EXAM == "10", "Ov", "NotF"))))))) %>% #comment if use Others category
#                              if_else(RES_EXAM == "9", "F+M", "Others")))))) %>%  
  group_by(RES_EXAM) %>%
  count() %>%
  mutate(pmType = n/5496673 * 100)
  

# Counting all cases by year
df %>%
  filter(RES_EXAM != "0") %>% 
  group_by(ANO) %>%
  count()

```

- With 2003: 5972921
### RES_EXAM categories, absolute number and percentages from 2003 to 2018:  
2 - Falciparum: 897411 (15.0 %)      
3 - F+FG: 151502  (2.54 %)      
4 - Vivax: 4782306 (80.1 %)      
5 - F+V: 44389  (0.743 %)     
6 - V+FG: 28042  (0.469 %)    
7 - FG: 46786  (0.783 %)      
8 - Malariae: 1928  (0.0323 %)       
9 - F+M: 577  (0.00966 %)      
10 – Ovale: 69 (0.00116 %)      
11 – Não F: 19911  (0.333 %)    

- Recategorization for paper description:  
Falciparum: 1095699 (18.3 %)      
Vivax: 4810348 (80.5 %)  
Falciparum + Malariae: 577  (0.00966 %)  
Falciparum + Vivax: 44389  (0.743 %)     
Malariae: 1928 (0.0323 %)   
Others (Ovale + Not Falciparum): 19980 (0.335 %)    


### RES_EXAM categories, absolute number and percentages from 2004 to 2018:  
F (Falciparum + F+FG + FG): 990280 (18.0 %)      
V (Vivax + V+FG): 4443459 (80.8 %)      
F+V: 40829  (0.743 %)     
Malariae: 1613  (0.0293 %)       
F+M: 538  (0.00979 %)      
Ovale: 43 (0.000782 %)      
Not F: 19911  (0.362 %) It can be all types except falciparum    

### Brazil 2018:  
```{r}
tmal18 <- df %>%
  filter(ANO == 2018) %>%
  mutate(RES_EXAM = if_else(RES_EXAM == "2" | RES_EXAM == "3" | RES_EXAM == "7" | RES_EXAM == "9", "F", 
          if_else(RES_EXAM == "4" | RES_EXAM == "6", "V",
                  if_else(RES_EXAM == "5", "F+V", "Others")))) %>%   
  group_by(RES_EXAM) %>% 
  count() %>% 
  mutate(pmType = n/242797 * 100)

```

- 2018 description without and with double counting mixed infection  
cases: 242,797    
V: 216769 (89.3 %), 217685 (89.7 %)    
F: 20357  (8.38 %), 21273 (8.8%)     


### Amazon Basin top state case per year:  
```{r}
df %>%
  filter(RES_EXAM != "0") %>% 
  select(UF_NOTIF, RES_EXAM, ANO) %>%
  group_by(ANO, UF_NOTIF) %>% 
  count() %>%
  filter(ANO == 2018) %>% #change the year
  arrange(desc(n))

```

- Top states per year  
2003 to 2008: 13    
2009 to 2012: 15  
2013 to 2018: 13 


### Cases per year only vivax and falci- double counting for mixed infection:  
2 - Falciparum: 
3 - F+FG: 151502  (2.54 %)      
4 - Vivax: 4782306 (80.1 %)      
5 - F+V: 44389  (0.743 %)     
6 - V+FG: 28042  (0.469 %)    
7 - FG: 46786  (0.783 %)      
8 - Malariae: 1928  (0.0323 %)       
9 - F+M: 577  (0.00966 %)      
10 – Ovale: 69 (0.00116 %)      
11 – Não F: 19911  (0.333 %)    

```{r}
## Vivax (vivax + mixed + vivax+FG)
vyear <- df %>%
  filter(RES_EXAM == "4" | RES_EXAM == "5" | RES_EXAM == "6") %>%   
  group_by(ANO) %>%
  count()
## Falci (falciparum + falciparum+FG + mixed + FG + mixedF+M)
fyear <- df %>%
  filter(RES_EXAM == "2" | RES_EXAM == "3" | RES_EXAM == "5" | RES_EXAM == "7" | RES_EXAM == "9") %>%   
  group_by(ANO) %>%
  count()

```

## Vivax, falciparum and total API and top API states per year

### Amazon Basin Total API per year
```{r}
# Amazon Basin cases per year
cases_AB_year <- df %>%
    filter(RES_EXAM != "0") %>%
    mutate(ANO = as.integer(ANO)) %>%
    group_by(ANO) %>%
    count() %>%
    rename(CASES = n)

# Amazon Basin population size per year
pop_AB_year <- pop %>%
  filter(CODE == 11 | CODE == 12 | CODE == 13 | CODE == 14 | CODE == 15 | CODE == 16 | CODE == 17 | CODE == 21 | CODE == 51) %>%
  select(-LEVEL, -NAME, -GR) %>%
  gather(key = ANO, value = POP_SIZE, -c(CODE)) %>% 
  mutate(ANO = as.integer(ANO)) %>% 
  group_by(ANO) %>% 
  summarise(POP_SIZE = sum(POP_SIZE))

# Amazon Basin API  
API_AB_year <- left_join(cases_AB_year, pop_AB_year) %>%
  mutate(API = CASES/POP_SIZE * 1000)

```

### Amazon Basin vivax and falciparum API per year
```{r}
# Amazon Basin cases per year
cases_AB_year <- df %>%
    filter(RES_EXAM != "0") %>%
    mutate(ANO = as.integer(ANO)) %>%
    group_by(ANO) %>%
    count() %>%
    rename(CASES = n)

vyear <- df %>%
  filter(RES_EXAM == "4" | RES_EXAM == "5" | RES_EXAM == "6") %>%       
  mutate(ANO = as.integer(ANO)) %>% 
  group_by(ANO) %>%
  count() %>% 
  rename(VIVAX = n) 

fyear <- df %>%
  filter(RES_EXAM == "2" | RES_EXAM == "3" | RES_EXAM == "5" | RES_EXAM == "7" | RES_EXAM == "9") %>%    
  mutate(ANO = as.integer(ANO)) %>% 
  group_by(ANO) %>%
  count() %>% 
  rename(FALCIPARUM = n) 


# Amazon Basin vivax and falciparum API  
APIvf_AB_year <- left_join(vyear, fyear)
APIvf_AB_year <- left_join(APIvf_AB_year, pop_AB_year)%>%
  mutate(APIV = VIVAX/POP_SIZE * 1000,
         APIF = FALCIPARUM/POP_SIZE * 1000)

```

### Top state API
```{r}
# Amazon Basin cases per year
cases_AB_yearTOP <- df %>%
    filter(RES_EXAM != "0") %>%
    mutate(ANO = as.integer(ANO),
           UF_NOTIF = as.integer(UF_NOTIF)) %>%
    group_by(ANO, UF_NOTIF) %>%
    count() %>%
    rename(CASES = n,
           CODE = UF_NOTIF) 

# Amazon Basin population size per year
pop_AB_yearTOP <- pop %>%
  filter(CODE == 11 | CODE == 12 | CODE == 13 | CODE == 14 | CODE == 15 | CODE == 16 | CODE == 17 | CODE == 21 | CODE == 51) %>%
  select(-LEVEL, -NAME, -GR) %>%
  gather(key = ANO, value = POP_SIZE, -c(CODE)) %>% 
  mutate(ANO = as.integer(ANO)) 

# Top API states per year
API_top <- left_join(cases_AB_yearTOP, pop_AB_yearTOP) %>% 
  mutate(API = CASES/POP_SIZE * 1000)

API_top %>% 
  filter(ANO == 2018) %>% #change the year
  arrange(desc(API))

```


