---
title: "SIVEP dictionary and cleaning"
author: "Raquel Lana"
date: "5 de dezembro de 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

# SIVEP - Epidemiological Surveillance Information System 
This file was done to share details of SIVEP variables and register all dropped variables and modifications we did on kept variables.  
  
## SIVEP time and space resolution  

SIVEP started in 2003 January. We have complete data from 2003 to 2018. The database has entries, permitting to aggregate into **daily case numbers** (reported date in year/month/day) and **epidemiological week**. These can later be aggregated to a different temporal resolution if needed.   

### Spatial resolution  
SIVEP is database for malaria notification that is mandatory in Brazil. Before 2003, the database was named SISMAL, which was a different database in relation to variables and categories.  

Each notification is associated with a federal unity, municipality, zone type (urban/rural) and a locality, which gives a finer resolution. Localities might be on the size of census tract or, more generally, smaller. Information from localities can be linked to IBGE census. SIVEP database has a complementary databe with localities information. This information will be listed below.

## Confirmation of malaria cases
Cases in Brazil are mostly confirmed by microscopy. The detection is of three types: passive, active and LVC (Cure Verification Thick Smear). **Passive case detection** corresponds to a diagnosis done when an individual looks for health assistance. **Active case detection** corresponds to a diagnosis done by a screening after finding a case. The protocol is described as an active search, when a case is reported, such that the neighboorhood around 1 km radius is investigated. LVC cases are notifications done when individuals return to health units. We still need to investigate more about numbers of LVC cases.  

## Treatment data     
SIVEP provides the **treatment schema for reported cases**. Completeness is > 85% and after 2011, > 95%. We have just requested more information about it, because is in a complementary database. 


## Variables book

### Kept variables

- **COD_NOTI**: notification number. 
- **DT_NOTIF**: date of notification (yyyy-mm-dd).    
- **TIPO_LAM**: type of detection (PCD/ACD/LVC).  
- **UF_NOTI**: state of notification.   
- **MUN_NOTI**: municipality of notification.  
- **DT_NASCI**: date of birth (yyyy-mm-dd).      
- **ID_PACIE**: numeric variable for age.    
- **ID_DIMEA**: if the age is per days, months and years.    
- **SEXO**: gender (Male/Female/Ignored).      
- **PAIS_RES**: informs the country code referring to the address of the patient according to IBGE.     
- **UF_RESID**: informs the state code referring to the address of the patient according to IBGE.  
- **MUN_RESI**: informs the municipality code referring to the address of the patient according to IBGE.  
- **LOC_RESI**: informs the locality code of residence according to the complementary database.  
- **SINTOMAS**: informs whether or not the patient had malaria symptoms (Yes/No).  
- **DT_SINTO**: informs the date when the patient felt the first symptoms of malaria (yyyy-mm-dd).  
- **COD_OCUP**: code of the main activity performed by the patient in the last 15 days (agriculture/livestock/domestic/turism/artisanal and small mining/exploration plant/hunting and fishing/Construction of Roads and Dams/mining/"traveler"/Others/Ignored).  
- **UF_INFEC**: informs the probable state of infection according to the code provided by IBGE.  
- **PAIS_INF**: informs the probable country of infection according to the code provided by IBGE.  
- **MUN_INFE**: informs the probable municipality of infection according to the code provided by IBGE.
- **LOC_INFE**: informs the probable locality of infection according to the complementary database.  
- **DT_EXAME**: date the exam was performed.  
- **RES_EXAM**: result of the exame stratified by Negativa (negative exam), Falciparum, F+FG, Vivax, F+V, V+FG, FG, Malariae, F+M, Ovale, Não F.
- **QTD_PARA**: reports the number of parasites per mm3.  
- **QTD_CRUZ**: it reports the amount of parasitemia in crosses, according to the following table: "< +/2" (minor than 1/2 cross), "+/2" (1/2 cross), "+" (one cross), "++" (two crosses), "+++" (three crosses), "++++" (four crosses).  
- **DT_TRATA**: informs date of start of treatment (FORMATO).
- **ID_LVC**: informs when notification is a case of LVC (Notification is a case of LVC/Notification is not a case of LVC).
- **GESTANTE_**: informs if the patient is pregnant (1st Quarter/2nd Quarter/3rd Quarter/Ignored gestational age/No/Not applicable).  
- **NIV_ESCO**: informs the level of education of the patient (Illiterate/1st to 4th incomplete series of Elementary School/4th complete series of Elementary School/5th to 8th grade incomplete elementary school/Complete primary school/incomplete high school/Complete high school/Incomplete higher education/Complete higher education/Not applicable).  
- **ESQUEMA**: informs the code of the treatment scheme used based on a complementary database. **We obtained all the information. See below**.
**COD_UNIN**: informs the code of the Notifying Unit.

### Dropped variables
- **DT_ENVLO**: reports the date the lot of notifications was received in the national data file (yyyy-mm-dd).
- **DT_DIGIT**: informs the date the notification was entered (yyyy-mm-dd).  
- **SEM_NOTI**: informs the date of notification in epidemiological week (SE/YEAR). Dropped because I have been extracted the epidemiological week and year before.
- **COD_AGEN**: informs the code of the agent who made the notification of the exam.  
- **RACA**: race/color (White/Black/Yellow/Brown/Indigenous).  
- **GESTANTE1**: pregnant (Yes/Not/Not applicable).  
- **NIV_ESCO_1**: schooling in years.
- **ESQUEMA_1**: Informs the treatment scheme used, based on old database.
- **HEMOPARASI**: reports the results of the examination for other hemoparasites screened according to the table below (Negative/Trypanosoma sp./Microfilaria/Trypanosoma sp. + Microfilariae/Not searched).  
- **EXAME**: informs the type of examination carried out (thick smear/RPD).  
- **EXAMINADOR**: inform the code of the professional who took the exam.  
- **VIVAX**: informs if the patient received treatment for vivax malaria in the last 60 days prior to notification.  
- **FALCIPARUM**: informs if the patient received treatment for falciparum malaria in the last 40 days prior to notification.

*Obs.: before here, the database was grouped and saved in another script outside Github.*  

- Loading RData file and libraries
```{r}
library(foreign)
library('stringr')
library("tidyverse")
library(lubridate)
library(data.table)
library(bit64)
require(maptools, quietly = TRUE, warn.conflicts = FALSE)
library('maptools')
library(rgdal)

#load("~/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/RData_full.RData") #until first weeks of 2018
load("/home/rlana/Dados/RData_full_new.RData") #until the end of 2018 and a update of 2017.

```

### Removing variables  
**Obs.**: RES_EXAM = 0 happens only in 2003 in 17 entries. I removed because this category does not exist.  
- 5972938 notifications from 2003 to 2018. 5972938 - 17 wrong entries in 2003 = 5972921

```{r}
#df <- df %>% 
      # select(-DT_ENVLO, -DT_DIGIT, -SEM_NOTI, -COD_UNIN, -COD_AGEN, -RACA, -GESTANTE1, -NIV_ESCO_1, -ESQUEMA_1, -HEMOPARASI, -EXAME_, -EXAMINADOR, -VIVAX, -FALCIPARUM) %>% 
      # filter(RES_EXAM == "2" | RES_EXAM == "3" | RES_EXAM == "4" | RES_EXAM == "5" | RES_EXAM == "6" | RES_EXAM == "7" | RES_EXAM == "9")

df <- df %>% 
      select(-DT_ENVLO, -DT_DIGIT, -SEM_NOTI, -COD_AGEN, -GESTANTE1, -NIV_ESCO_1, -ESQUEMA_1, -HEMOPARASI, -EXAME_, -EXAMINADOR, -VIVAX, -FALCIPARUM) %>% 
      filter(RES_EXAM == "2" | RES_EXAM == "3" | RES_EXAM == "4" | RES_EXAM == "5" | RES_EXAM == "6" | RES_EXAM == "7" | RES_EXAM == "9")

```

### Filtering and grouping to falciparum and vivax types.
RES_EXAM categories, absolute number and percentages from 2003 to 2018:  
2 - Falciparum: 897411 (15.0 %)      
3 - F+FG: 151502  (2.54 %)      
4 - Vivax: 4782306 (80.1 %)      
5 - F+V: 44389  (0.743 %)     
6 - V+FG: 28042  (0.469 %)    
7 - FG: 46786  (0.783 %)      
8 - Malariae: 1928  (0.0323 %)       
9 - F+M: 577  (0.00966 %)      
10 – Ovale: 69 (0.00116 %)      
11 – Não F: 19911  (0.333 %)    


- Categorizing variables and transforming to factor.
```{r}
df <- df %>%
  mutate(RES_EXAM = as.factor(case_when(RES_EXAM == "2" | RES_EXAM == "3" | RES_EXAM == "7" | RES_EXAM == "9" ~ "Falciparum",
                                        RES_EXAM == "4" | RES_EXAM == "6" ~ "Vivax",
                                        TRUE  ~  "Mixed"))) %>%
  mutate(SINTOMAS = as.factor(case_when(SINTOMAS == "1" ~ "YES", 
                                        TRUE ~ "NO"))) %>% 
  mutate(QTD_CRUZ = as.factor(case_when(QTD_CRUZ == "1" ~ "< +/2",
                                        QTD_CRUZ == "2" ~ "+/2",
                                        QTD_CRUZ == "3" ~ "+",
                                        QTD_CRUZ == "4" ~ "++",
                                        QTD_CRUZ == "5" ~ "+++",
                                        QTD_CRUZ == "6" ~ "++++",
                                        TRUE ~ "NA"))) %>% 
  mutate(ID_LVC_ = as.factor(case_when(ID_LVC_ == "1" ~ "YES",
                                       TRUE ~ "NO"))) %>% 
  mutate(GESTANTE_ = as.factor(case_when(GESTANTE_ == "1" ~ "1qtr",
                                         GESTANTE_ == "2" ~ "2qtr",
                                         GESTANTE_ == "3" ~ "3qtr",
                                         GESTANTE_ == "4" ~ "Ignored",
                                         GESTANTE_ == "5" ~ "NO",
                                         GESTANTE_ == "6" ~ "NA",
                                         TRUE ~ "NA"))) %>% 
  mutate(TIPO_LAM = as.factor(case_when(TIPO_LAM == "1" ~ "PCD",
                                        TIPO_LAM == "2" ~ "ACD",
                                        TRUE ~ "CV"))) %>% 
  mutate(SEXO = as.factor(SEXO),
         COD_OCUP = as.factor(COD_OCUP),
         NIV_ESCO = as.factor(NIV_ESCO))

#levels(df$SINTOMAS)

```

## Adding AGE RANGE

### Creating the continuos age by year and after by categories (age range) from IBGE
- Continuos age in year: here I dropped registries with inconsistency.
```{r}
df <- df %>% 
  filter(ID_PACIE < 30 & ID_DIMEA == "D" | ID_PACIE < 12 & ID_DIMEA == "M" | ID_PACIE <= 100 & ID_DIMEA == "A") %>% 
  mutate(ID_PACIE = as.numeric(ID_PACIE)) %>% 
  mutate(AGE_CONT = as.numeric(if_else(ID_DIMEA == "A", ID_PACIE, 0)))

```

- Age range
```{r}
labs <- c("0-4","5-9","10-14","15-19","20-24","25-29","30-34","35-39","40-44","45-49","50-54","55-59","60-64","65-69","70-74","75-79","80+")
df <- df %>% 
  mutate(AGE_CAT = cut(AGE_CONT, breaks = c(seq(0, 80, by = 5), Inf), labels = labs, right = FALSE))
           
```


- Convert some characters to integer
```{r}
df <- df %>% 
  mutate(COD_NOTI = as.integer(COD_NOTI),
         UF_NOTIF = as.integer(UF_NOTIF),
         MUN_NOTI = as.integer(MUN_NOTI),
         COD_UNIN = as.integer(COD_UNIN),
         PAIS_RES = as.integer(PAIS_RES),
         UF_RESID = as.integer(UF_RESID),
         MUN_RESI = as.integer(MUN_RESI),
         LOC_RESI = as.integer(LOC_RESI),
         UF_INFEC = as.integer(UF_INFEC),
         PAIS_INF = as.integer(PAIS_INF),
         MUN_INFE = as.integer(MUN_INFE),
         LOC_INFE = as.integer(LOC_INFE))

```



Saving RData.
```{r}
save(df, file = "/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Malaria_Mapping_TimeSeries_Data/SIVEP_clean_v1.RData")
save(df, file = "/home/rlana/Dados/SIVEP_clean_2003_2018.RData")

```


## Schema treatment database
- Reading the schema treatment database
```{r}
envNN = F

if(envNN){
  treat <- read.dbf("/Malaria_Mapping_TimeSeries_Data/ESQUEMA_.DBF", as.is = F)
}else{treat <- read.dbf("/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Malaria_Mapping_TimeSeries_Data/ESQUEMA_.DBF", as.is = F)
}

```

- Categories
```{r}
unique(treat$CD_ESQUEMA)
unique(treat$ESQUEMA)
unique(treat$STATUS)
treat %>% filter(CD_ESQUEMA == 12)

```

#### Details about treatment schema
This variable is not explored here. Go to *Manipulate_data_maps_ts.rmd*.    
**ESQUEMA**  
1- Infections by *P. vivax* or *P.ovale* with Chloroquine in 3 days and Primaquine in 7 days (short scheme).    
2- Infections by *P. vivax* or *P.ovale* with Chloroquine in 3 days and Primaquine in 14 days (long scheme).    
3- Infections for *P. malariae* for all ages and for *P. vivax* or *P. ovale* in pregnant women and children under 6 months, with Chloroquine in 3 days.    
4- Prevention of frequent relapses by *P. vivax* or *P. ovale* with weekly chloroquine at 12 weeks.    
5- *P. falciparum* infections with fixed combination of Artemeter + Lumefantrine in 3 days.    
6- *P. falciparum* infections with fixed combination of Artesunate + Mefloquine in 3 days.    
7- *P. falciparum* infections with Quinine at 3 days, Doxycycline at 5 days and Primaquine at 6 days.    
8- Mixed infections by *P. falciparum* and *P. vivax* or *P.ovale* with Artemeter + Lumefantrine or Artesunate + Mefloquine in 3 days and Primaquina in 7 days.    
9- Uncomplicated infections by *P. falciparum* in the first trimester of gestation and children under 6 months, with quinine in 3 days and Clindamycin in 5 days.    
10- Severe and complicated malaria by *P. falciparum* in all age groups.    
11- *P. falciparum* infections with fixed combination of Artemeter + Lumefantrine in 3 days and primaquine in a single dose.      
12- *P. falciparum* infections with fixed combination of Artemeter + Mefloquine in 3 days and primaquine in a single dose.      
83- Mixed infections by *P. vivax* + *P. falciparum* with Mefloquine in single dose and Primaquine in 7 days.  
85- *P. vivax* infections in children with vomiting, with rectal capsules of Artesunate in 4 days and Primaquina in 7 days.  
86- Infections of *P. falciparum* with Mefloquina in single dose and Primaquina in the second day.  
87- *P. falciparum* infections with Quinine at 7 days.  
88- *P. falciparum* infections in children with Artesunate rectal capsules in 4 days and single dose of Mefloquine in the 3rd day and Primaquina in the 5th day.  
89- Mixed infections with *P. vivax* + *P. falciparum* with Quinine at 3 days, Doxycycline at 5 days and Primaquine at 7 days.  
99- Another scheme used (by doctor).    

- **Obs.**: 83, 85, 86, 87, 88, 89 are out of use now.


**ESQUEMA_1** (Field with codes in the template of the old notification sheet) - **dropped**.
1- Pv with Chloroquine Infections in 3 days and Primaquine 7 days  
2- Pf infections with Quinine in 3 days + Doxycycline in 5 days + Primaquine in the 6th day  
3- Mixed Pv + Pf Infections with Mefloquine in a single dose and Primaquina in 7 days  
4- Pm Chloroquine Infections in 3 Days  
5- Pv infections in children with vomiting, with artesunate rectal capsules in 4 days and Primaquina in 7 days  
6- Infections by Pf with Mefloquina in single dose and Primaquina in the second day  
7- Pf Infections with Quinine at 7 Day  
8- Pf Infections of children with rectal artesunate capsules in 4 days and single dose of Mefloquine on day 3 and Primaquine on day 5  
9- Mixed infections by Pv + Pf with Quinine in 3 days, Doxycycline in 5 days and Primaquina in 7 days  
10- Prevention of Malaria Relapse by Pv with Chloroquine in a single weekly dose for 3 months  
11- Severe and complicated malaria  
99- Other Scheme used (per physician)  



### Localities data availability in SIVEP

**CD_LOCAL**: informs the code of the locality according to the geographical recognition of SMS.        
**NM_LOCAL**: informs the name of the locality according to the geographical recognition of SMS.        
**UF_IBGE**: informs the code corresponding to the UF (State) where it is located.  
**MUN_IBGE**: inform the code of the municipality where the locality is located.   
**NU_PREDI**: informs the number of existing buildings in the locality. **This SIVEP information is out of date for several locations.**     
**NU_HABIT**: informs the number of inhabitants in the locality. **This SIVEP information is out of date for several locations.**        
**DT_REG**: informs the date when the locality was registered.      
**DT_ATUAL**: informs the date in which the number of inhabitants was updated, number of buildings, among others, of the locality.   
**LONGITUD**: informs the coordinate referring to the longitude where the geographical center of the locality is located.   
**LATITUDE**: informs the coordinate about the latitude where the geographical center of the locality is located. 
**ZONA**: informs the Zone (urban / rural) where the locality is located, 1-Urban and 2-Rural.      
**CATEGORI**: informs the category to which the locality belongs, according to the table of categories attached:           
1- SITIO (SIT)  
2- FAZENDA (FAZ.)  
3- BAIRRO (BAIR)  
4- POVOADO (POVO)    
5- VILA (VILA)    
6- NUCLEO (NUCL)    
7- CIDADE/SEDE MUNICIPAL (CID.)    
8- USINA (USIN)    
9- CONJUNTO HABITACIONAL (CONJ)    
10- ENGENHO (ENGE)   
11- ACAMPAMENTO (ACAM)   
12- MALOCA (MALO)    
13- SERRARIA (SERR)    
14- QUADRA (QUAD)    
15- QUARTEIRAO (QUAR)    
16- BLOCO DE APARTAMENTO (BLOC)    
17- GARIMPO (GARI)    
18- CEMITERIO (CEMI)    
19- CASA ISOLADA (CASA)    
20- PRACA/PARQUE (PRAC)    
21- IGREJA/TEMPLO (IGRE)      
22- AEROPORTO (AERO)    
23- ESTADIOS/GINASIOS (ESTA)    
24- CLUBES (CLUB)    
25- RODOVIARIA/FERROVIARIA (RODO)    
26- COLONIA (COL)    
27- SERINGAL (SER)    
28- GLEBA (GLE)    
29- ALDEIA (ALD)    
30- RAMAL (RAM)    
31- PROJETO (PROJ)    
32- PROJETO ASSENTAMENTO DIRIGIDO (PAD)    
33- BAIXAO (BAIX)    
34- GROTA (GROT)    
35- LINHA (LINH)    
36- ILHA (FAVE)    
37- FAVELA (FAVE)    
38- CORTICO (CORT)    
39- MORRO (MORR)    
40- LAGOA (LAGO)     
41- ESTRADA (ESTR)    
42- PORTO (PORT)    
43- AREA DE ABRANGENCIA (ABRG)    

**AGLOMERA**: informs the code of the agglomerate to which the locality belongs.   
**STATUS**: informs the status of the locality, 1-Active and 2-Extinct.   
**CODSEI**: ?   
**COPOBASE**: ?
```{r}
l1 <- read.dbf("/home/claudia/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Dados/LOCANOT.DBF")
l1 <- read.dbf("/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Dados/LOCANOT.DBF", as.is = F)

head(l1)
str(l1)
tail(l1)

write.csv(l1, file = "/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Dados/locality_database.csv", row.names = F)

l1 <- fread("/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Dados/locality_database.csv", encoding = "UTF-8")

l1 <- fread("/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Dados/locality_database.csv", encoding = "Latin-1")

```


### Localities data availabilty in IBGE
- Example of Acre state  
IBGE database has information for census tract, zone (urban or rural), macro, meso and micro regions, municipality and state. The finest resolution to calculate the API is for zone. For locality we have to use population size provided by SIVEP, that can be very outdated depends on the municipality.    
```{r}
l2 <- fread("/home/claudia/Documentos/Doutorado/Desenvolvimento/Dados/Shapes/IBGE/Base_informacoes_setores2010_universo_AC/Base informaçoes setores2010 universo AC/CSV/Basico_AC.csv", sep2 = ";", encoding = "Latin-1", stringsAsFactors = F)
head(l2)
class(l2$Cod_setor)
unique(l2$Cod_setor)

l2$Cod_setor <- as.integer(l2$Cod_setor)

```


## Settlement data, minimg, dams and developments in general

### Acre as an example
Municipality limity shapefile.
```{r}
ibge <- readOGR("/home/claudia/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Dados/AC_IBGE", layer = "12MUE250GC_SIR")
plot(ibge)

```


#### Settlements
Land collection of the National Institute of Colonization and Agrarian Reform (Instituto Nacional de Colonização e Reforma Agrária). State, name of the settlement project, municipality, are, capacity, number of families, fase, date of creation, way of obtaining, date of obtaining, description.    
site: http://acervofundiario.incra.gov.br/acervo/acv.php
```{r}
#mapa <- readOGR("/home/raquel/Documentos/PosDoc_PROCC/OwncloudPD/malaria/AnalisesRaquel/Pasteur/Dados/Assentamento Brasil_AC/", layer="Assentamento Brasil_AC")
mapa <- readOGR("/home/claudia/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Dados/Assentamento Brasil_AC", layer="Assentamento Brasil_AC")
plot(mapa, col = "red", add = T)
mapa@data
head(mapa@data)

```

### Mining
The site allows to download the information by state and has all kinds of process that arrive for the National Department of Mineral Production (Departamento Nacional de Produção Mineral). In this database has the number of process, year that began, area, fase, las event, name of the mining, mineral explored, use and state.          
site: http://www.anm.gov.br/assuntos/ao-minerador/sigmine
```{r}
mapa1 <- readOGR("/home/claudia/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Dados/Mineracao_AC/AC", layer = "AC")
plot(mapa1, col = "blue", add = T)
head(mapa1@data)
mapa1_proj <- spTransform(mapa1, CRS("+proj=longlat +ellps=GRS80 +no_defs")) 
proj4string(mapa1_proj)
plot(mapa1_proj, col = "blue", add = T)

stopifnot(proj4string(mapa) == proj4string(mapa1))
stopifnot(proj4string(mapa) == proj4string(ibge))
stopifnot(proj4string(mapa1) == proj4string(ibge))

```

### Dams
The National Water Agency (Agência Nacional de Águas) has the National Information System on Dams Safety (Sistema Nacional de Informações sobre Segurança de Barragens) that provides information about the localtion, status, names, etc about dams in an Excell format.
site: http://www.snisb.gov.br/portal/snisb/mapas-tematicos-e-relatorios/tema-1-1


### Hydroeletric plant
The National Electric Energy Agency (Agência Nacional de Energia Elétrica) has the Geographic Information System of the Electric Sector (Sistema de Informações Geográficas do Setor Elétrico) that provides shapefiles for different types of energy eletric developments and their locations, status, etc.
site: https://sigel.aneel.gov.br/Down/

```{r}
mapa2 <- readOGR("/home/claudia/Documentos/PosDoc_PROCC/malaria/AnalisesRaquel/Pasteur/Dados/Barragens_Usinas/zipfolder", layer = "Usinas_Hidreletricas_UHE")
proj4string(mapa2)
mapa2_proj <- spTransform(mapa2, CRS("+proj=longlat +ellps=GRS80 +no_defs")) 
plot(mapa2_proj)
head(mapa2_proj@data)

```


### Variables doubts list
1- Severe malaria: what means? About medicines, hospitalizations?  
2- Pregnant women using Primaquine  
3- TIPO_LAM and LVC- why LVC is decreasing during the years? What's the criteria to fill?  
4- How they fill date of 1st symptoms, date of exam and treatment: many problems about the difference in days.   
5- What is CODSEI and COPOBASE from locality database? I didn't find on the locality database dictionary.  
6- What is SMS acronym founded in some explanations in loc database?
7- The variable about occupation has the category "Others" that has an expressive number of the registry. How is this category filled in? What kind of occupation is included in this category?
